---
title: "Distribution of genetic paternity among male primates"
author: "Stacy Rosenbaum, Nicholas Grebe, Joan B. Silk"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: references.bib
nocite: |
  @Nievergeltetal2002
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = file.path(dirname(input), "..", "output"),
      ...
    )
  })
---

# Abstract

<!--Reminder to figure out the logistics of calculating the numbers in the abstract up front so that they can be fully automated-->

Understanding the distribution of paternity within social groups is critical for testing hypotheses about the evolution of behavior and morphology in primates, but assembling the requisite comparative data is a challenging task. We compiled genetic paternity data from 52 species of wild nonhuman primates along with information about ecological, morphological, and social traits that are potentially relevant to understanding what proportion of offspring are sired by "primary males" (i.e., alpha males in multi-male groups and resident males in single male groups). Our data, which currently contains information about 11 primate Families and >3,000 individual paternities, is presented as a publicly accessible, living database designed to be updated as new data become available. Using Bayesian phylogenetic Beta regression and binomial models, we answered questions about the role that phylogeny, group composition, and seasonality play in determining primary males' paternity share, as well as determined the relative share of paternities obtained by non-primary residents versus extra-group males. First, we found that phylogeny has a detectable but relatively modest influence on primary males' paternity share. Species-level differences explained roughly 35–40% of variation in primary males' paternity share, and of that interspecific variation, ~50-70% was attributable to shared phylogenetic history. Second, group composition strongly predicted paternity outcomes. Primary males in single-male/multi-female groups obtained the highest share of paternity, while those in multi-male groups had the lowest, though there was substantial variation within each category. Pair-living animals showed a striking split: males in cohesive pairs sired ~90% of offspring, while those in dispersed pairs sired only ~55%. Contrary to theoretical expectations, reproductive seasonality did not predict primary males' paternity share in any group type. Finally, when primary males in multi-male groups lost paternities, ~75% of losses were to other resident males. Overall, ~5-15% of offspring were sired by extra-group males. Our results largely confirm earlier findings based on smaller data sets, but also that the relationship between social organization and paternity is more complicated than simple categorical predictions suggest. We discuss the gap between the data that would ideally be available for testing these hypotheses versus what currently exists, with hopes that our living database can help close this gap over time.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache=FALSE)
library(dplyr)
library(readxl)
library(gt)
library(brms)
library(tidybayes)
library(tidyr)
library(ggplot2)
library(ape)
library(phytools)
library(ggtree)
library(patchwork)
library(posterior)
library(stringr)

# Clean and merge the raw Excel data (replaces the former Stata pipeline)
source("clean_and_merge_data.R")

# Auto-numbering counters for tables and figures
table_count <- 0
table_num <- function() {
  table_count <<- table_count + 1
  return(table_count)
}

fig_count <- 0
fig_num <- function() {
  fig_count <<- fig_count + 1
  return(fig_count)
}

# Pre-compute table numbers for cross-referencing in body text
# Make sure that these match the order in which table_num() is called when tables are created
tbl_studyid_num <- table_count + 1       
tbl_context_num <- table_count + 2       
tbl_taxonomy_num <- table_count + 3      
tbl_paternity_num <- table_count + 4     
tbl_traits_num <- table_count + 5        
tbl_exclusion_num <- table_count + 6     
tbl_grptype_num <- table_count + 7       

# Insert figure captions so they can be referenced in chunk headers
fig_studies <- fig_count + 1
fig_studies_num <- fig_num()  # increment counter here, but define caption text later (needs data)

fig_simulation <- fig_count + 1
fig_simulation_cap <- paste0("Figure ", fig_num(), ". Expected change in the proportion of paternities obtained by primary males as a function of number of competitors and fertile female:male ratio. E.g., a 5:1 ratio (the blue line) would represent a situation where a large number of females were fertile at the same time, and thus monopolizability is low\u2013for example, in a seasonally breeding monkey species. A 1:5 ratio (represented by the purple line) would represent a situation where individual females only occasionally come into estrus (e.g., a chimpanzee group that contains many males but likely only one or at most two fertile females at any given point in time), and so monopolizability is high.")

fig_tree <- fig_count + 1  
fig_tree_num <- fig_num()  # increment counter here, but define caption text later (needs data)

fig_patloss <- fig_count + 1
fig_patloss_cap <- paste0("Figure ", fig_num(), ". Males in multi-male (MM) groups had the lowest share of paternity in the raw data (the small circles) and in model-derived predicted effects (black bars). The credible interval for the multi-male to single-male/multi-female (MFF) comparison did not include zero. Therefore, we can be confident that primary males in single-male/multi-female groups obtain more paternities than males in multi-male groups. The credible intervals for the other two comparisons (i.e., multi-male to pair-living (MF), and pair-living to single-male/multi-female) included zero. The raw data mean for pair-living species where pairs are cohesive is represented by a green diamond, while the mean for pair-living species where pairs are dispersed is represented by an orange diamond. The n=2 observations where multi-male groups contained multiple males and only a single female are represented in blue. ")

fig_seasonality <- fig_count + 1
fig_seasonality_cap <- paste0("Figure ", fig_num(), ". Seasonality had little effect on primary male paternity share. In both multi-male and single-male groups, species that were least seasonal, moderately seasonal, and most seasonal had similar estimated primary male paternity proportions, with extensive overlap in credible intervals. The results were qualitatively very similar when pair-living species were excluded from the single-male category. ")

```

```{r caption-style, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<style>\n  .figure-caption, .caption { font-size: 0.8em; }\n</style>\n')
}
```

```{r load-prepare-data, echo=FALSE, results='hide'}

paternity_table <- read.csv("../data/paternity-table/paternity_table_merged.csv") %>%
  mutate(
    proportion_egp_minimum = n_egp / n_paternities,
    proportion_primary_minimum = n_alphaprimary / n_paternities
  ) %>%
  rowwise() %>%
  mutate(
    n_egp_maximum = sum(n_egp, n_egp_ambiguous, na.rm = TRUE), #counts NAs as 0s
    n_primary_maximum = case_when(
      !is.na(n_alphaprimary) ~ sum(n_alphaprimary, n_residents_unknownrank, na.rm = TRUE),
      TRUE ~ NA),  # NAs for alpha always return NAs
    proportion_primary_maximum = n_primary_maximum / n_paternities,
    proportion_egp_maximum = n_egp_maximum / n_paternities
  ) %>%
  ungroup() %>%
  mutate(
    grp_composition = as.factor(grp_composition),
    phylo = as.factor(genus_species_subspecies),
    spp = as.factor(genus_species_subspecies)
  ) %>%
  filter(!is.na(js_or_sr)) # getting rid of T. leucocephalus (no longer in data as of April 2024)

write.csv(paternity_table, "../data/paternity-table/paternity_table.csv")

data <- read.csv("../data/paternity-table/paternity_table.csv")
phylo_matrix <- as.matrix(read.csv("../data/phylogeny/phylo_matrix.csv", row.names = 1))
phylo_tree <- read.nexus("../data/phylogeny/consensusTree_10kTrees_Primates_Version3.nex")


# Make adjusted versions of the minimum and maximum variables to accommodate Beta regression
data <- data %>%
  mutate(
    proportion_primary_minimum_adj = case_when(
      proportion_primary_minimum == 0 ~ 0.001,
      proportion_primary_minimum == 1 ~ 0.999,
      TRUE ~ proportion_primary_minimum
    ),
    proportion_primary_maximum_adj = case_when(
      proportion_primary_maximum == 0 ~ 0.001,
      proportion_primary_maximum == 1 ~ 0.999,
      TRUE ~ proportion_primary_maximum
    ),
    proportion_egp_minimum_adj = case_when(
      proportion_egp_minimum == 0 ~ 0.001,
      proportion_egp_minimum == 1 ~ 0.999,
      TRUE ~ proportion_egp_minimum
    ),
    proportion_egp_maximum_adj = case_when(
      proportion_egp_maximum == 0 ~ 0.001,
      proportion_egp_maximum == 1 ~ 0.999,
      TRUE ~ proportion_egp_maximum
  )
)

# Make the species variable into a factor so it is usable as a random effect
data$genus_species_subspecies <- as.factor(data$genus_species_subspecies)

# Create a duplicate version of the species variable so it can be used alongside phylo matrix
data$species_nonphylo <- data$genus_species_subspecies

# Define fig_studies_cap here (rather than in setup) because it needs data-derived values
min_year <- min(paternity_table$publicationyear, na.rm = TRUE)
max_year <- max(paternity_table$publicationyear, na.rm = TRUE)
fig_studies_cap <- paste0("<span style='font-size: 85%;'>Figure ", fig_studies_num, ". Grey bars represent the number of studies published each year between ", min_year, " and ", max_year, " that incorporate paternity data from wild non-human primates. The orange line tracks the cumulative number of studies over time, while the blue line tracks the cumulative number of species represented across those studies. Data in our table that are classified as personal communication/unpub data are not included in this figure. Only one species, <i>Papio anubis</i>, was represented in unpublished data but not published data.")

```

# Introduction

Understanding the distribution of paternity within social groups or populations is critical for testing a wide variety of hypotheses about the evolutionary origins of behavior and morphology. For example, incomplete control models of male reproductive skew predict that the extent of reproductive skew within groups will decrease with increases in the degree of reproductive synchrony among females and the number of males in groups  [@Ostneretal2008]. The degree of sexual dimorphism that occurs in a given species is expected to be a product of the degree of male reproductive skew; in species where skew is high, sexual dimorphism in body size is greater and males are more likely to have ‘weaponry’ useful for fighting for access to females [@Andersson1994]. In species that typically live in pairs, male care and the strength of pair bonds are expected to be inversely related to rates of extra-pair paternity [@Hucketal2014]. And, the evolution of cooperative breeding in mammals is expected to be linked to high rates of maternal and paternal relatedness within groups [@LukasClutton-Brock2012].

Nonhuman primates are especially useful for exploring the drivers of paternity distribution because they are the second-largest mammalian Order and have extensive diversity in social organization and mating systems. The advent of DNA fingerprinting in the 1980s and the development of efficient methods for extracting DNA from fecal samples made it feasible to assess paternity in wild primate populations without capturing or handling them. The number of studies providing paternity data from wild nonhuman primates has steadily increased over time (Figure `r fig_studies`), and genetic paternity data are now available for more than 50 species. This represents a substantial increase over the number of nonhuman primate species that were available in previous reviews and comparative analyses [e.g., @Milleretal2021; @Clutton-BrockIsvaran2006; @GogartenKoenig2013; @Ostneretal2008; @Widdig2007; @Soulsbury2010]. In addition to increasing the total number of species available, researchers have accumulated larger samples of paternities for particular species and groups. Bigger sample sizes are particularly important if there is variation in patterns of paternity distribution across time within groups, or across populations within species; evidence from some nonhuman primate species suggests that this is frequently the case (Alberts, Watts, and Altmann 2003; Rosenbaum et al. 2015).

```{r fig-paternity-studies, echo=FALSE, out.width="80%", fig.cap=fig_studies_cap}

#This figure shows the accumulation of primate paternity data over time

# 1. Summarize data by year

# What is the first year each species appears?
first_appearances <- paternity_table %>%
  filter(!is.na(publicationyear)) %>%
  group_by(species_10ktrees) %>% #use this var so it doesn't count subspecies as separate species
  summarise(first_year = min(publicationyear), .groups = "drop") %>%
  count(first_year, name = "n_new_species")

# Summarize number of studies per year
temporal_trends <- paternity_table %>%
  filter(!is.na(publicationyear)) %>%
  group_by(publicationyear) %>%
  summarise(n_studies = n_distinct(reference), .groups = "drop") %>%
  arrange(publicationyear) %>%
  # Add species appearances and fill missing years
  left_join(first_appearances, by = c("publicationyear" = "first_year")) %>%
  replace_na(list(n_new_species = 0)) %>%
  # Calculate cumulative totals
  mutate(
    cumulative_studies = cumsum(n_studies),
    cumulative_species = cumsum(n_new_species)
  )

# Reality check
#tail(temporal_trends, 1)


# 2. Plotting

temptrends_plot <- ggplot(temporal_trends, aes(x = publicationyear)) +

  # Histogram of new studies (left y-axis)
  geom_bar(aes(y = n_studies), stat = "identity", fill = "grey90") +

  # Cumulative published studies (orange line, right y-axis)
  geom_line(aes(y = cumulative_studies / 10, color = "Cumulative published studies"), size = 1) +
  geom_point(aes(y = cumulative_studies / 10, color = "Cumulative published studies"), size = 2) +

  # Cumulative species (blue line, right y-axis)
  geom_line(aes(y = cumulative_species / 10, color = "Cumulative species represented"), size = 1) +
  geom_point(aes(y = cumulative_species / 10, color = "Cumulative species represented"), size = 2) +

  # Color mapping
  scale_color_manual(
    name = "",
    values = c("Cumulative published studies" = "darkorange",
               "Cumulative species represented" = "darkblue")
  ) +

  # Dual y-axes
  scale_y_continuous(
    name = "New studies per year",
    limits = c(0, 10),
    breaks = c(2, 4, 6, 8, 10),
    sec.axis = sec_axis(~ . * 10, name = "Cumulative count")
  ) +

  # Labels
  labs(
    x = "Year",
    title = "Number of studies incorporating\n paternity data from wild primates over time"
  ) +

  # Theme and text styling
  theme_classic() +
  theme(
    axis.title.y.right = element_text(color = "black", size = 16),
    axis.text.y.right  = element_text(color = "black", size = 14),
    axis.title.y.left  = element_text(size = 16),
    axis.title.x       = element_text(size = 16),
    axis.text          = element_text(size = 14),
    plot.title         = element_text(hjust = 0.5, size = 16),
    legend.position    = c(0.25, 0.85),
    legend.background  = element_blank(),
    legend.text        = element_text(size = 12),
    legend.key.size    = unit(0.6, "cm")
  )

print(temptrends_plot)

cumulative_species_fulltab <- length(unique(paternity_table$species_10ktrees)) # For inline text

```

Compiling information about paternity distribution and phenotypic traits from primary sources is a labor-intensive task, and thus it is extremely useful when comparative databases are compiled and made available to the academic community. Comparative analyses are more powerful when they include information from more species, and they are more accurate when measurements of traits are based on larger and more representative samples. This means it is important to include all available information, and to provide a mechanism for updating databases as new data become available.

Compiling information about paternity distribution and phenotypic traits from primary sources is a labor-intensive task, and thus it is extremely useful when comparative databases are compiled and made available to the academic community. Comparative analyses are more powerful when they include information from more species, and they are more accurate when measurements of traits are based on larger and more representative samples. This means it is important to include all available information, and to provide a mechanism for updating databases as new data become available.

We compiled information about the distribution of paternity in wild primate groups, along with complementary information about socioecological, morphological, and behavioral characteristics that are potentially useful for testing evolutionary hypotheses about paternity distribution. A critical goal of our project is to compile all available information on paternity distribution in wild primate groups, make these data freely available to the academic community, and to create a mechanism for updating the database as new data become available. We attempt to achieve three key objectives for comparative datasets: transparency, usability, and reproducibility [@Borriesetal2016]. 

Another immediate goal of our project is to use these data to answer three basic questions about paternity distribution in non-human primates: first, to what degree does phylogeny predict paternities obtained by alpha males in multi-male groups, and resident males in single-male groups? The answer to this question is useful for understanding the relative importance of evolutionary history versus current environmental conditions in determining paternity distribution. 

Our second question is, what is the magnitude of paternity loss for resident males? We compare if and how much this differs for males living in different types of groups. This is a test of the widely-accepted (and thus far well-supported) prediction that paternity loss should be a function of group composition, where pair-living (i.e., socially monogamous) males have the highest paternity certainty and those in multi-male/multi-female groups the least [@Clutton-BrockIsvaran2006]. We also investigate whether the proportion of paternities obtained is influenced by reproductive seasonality. This is a test of the prediction that males in species where female fertility is highly synchronized will have more difficulty monopolizing mating opportunities, which some prior metanalyses have found evidence for [e.g., @Paul1997; @Ostneretal2008] and some have not [e.g., @KutsukakeNunn2006]. Understanding the magnitude of group composition and seasonality effects is useful for understanding the selective pressures shaping male reproductive strategies. 

Third, we answer the question, when top-ranking males lose paternities in multi-male groups, who are they losing it to: other residents, or males outside their social group? This is useful for understanding the source of male-male competition, and the tradeoffs males make when they are either forced to, or choose to, live in social groups with other males. 

# Methods

## Data compilation

### Sources of information

During the dark days of COVID, JBS and SR began searching for primate paternity data in Google Scholar, using a variety of key words: paternity, male reproductive success, reproductive skew, mating system, inbreeding, kin discrimination, and kin recognition. We also tracked citations from lists previously compiled by others. In almost all cases, we obtained values from the original publications, not from secondary sources; the few exceptions are noted and full references provided in the database. For published abstracts, we contacted the lead authors and requested information about the distribution of paternity. In these cases, the abstract is cited along with a notation that more detailed data were obtained through personal communication with the author(s).  We also contacted field researchers that we thought might have unpublished paternity data and requested any information about paternity distribution that they were willing to share.

Some studies provide data from multiple social groups in the same population. When possible, we present information for each group separately. In some cases, these group-level data were available in the original publication. In other cases, we obtained information about individual groups from the authors of the original publication. In these cases, we cite the original source, along with a notation that additional data were obtained through personal communication with the authors. When we were unable to separate data from multiple groups, we compiled aggregate information and designated the group name as ‘multiple’ in the database.

In some instances, there is more than one study of the same group and the paternity samples overlap (e.g., the capuchin paternities in @Munizetal2010 are a subset of the paternities in @Godoyetal2016). In these cases, all the studies are included, and their overlap is indicated in the table, along with information about which one contains the largest sample size. To avoid duplication of data, our analyses are based on the study with the largest sample size.

### Types of information

Below we describe the data we aggregated, loosely subdivided into types of information for organizational purposes. We combined information from the paternity studies we used with information from other studies of the relevant species. The data are arranged so that each row corresponds to a particular social group (whenever possible) or to an aggregation of multiple social groups (when it was not possible to extract group-level statistics from a given publication).

#### Study identification

In order to identify individual studies and track entries over time, we include the reference to the paper the data came from, the year the paper was published, the data source (i.e., whether it was primary data or secondary), and the date we added the study to our table. These variables are listed in Table `r tbl_studyid_num`.

```{r study-identification-table, echo=FALSE}

# Variables to include
variables_of_interest <- c(
  "reference", 
  "publicationyear", 
  "datasource", 
  "dateofentry"
)

# Descriptions for each variable
descriptions <- c(
  reference = "Study the data came from",
  publicationyear = "Publication year",
  datasource = "Primary or secondary data source; if secondary, refs in notes",
  dateofentry = "Date that study was entered into the database"
)

# Build the table
variable_table <- tibble(
  `Variable name` = variables_of_interest,
  `Variable description` = descriptions[variables_of_interest]
) %>%
  gt() %>%
  tab_header(
    title = paste0("Table ", tbl_studyid_num, ". Study identification variables")
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size = px(13),              
    table.align = "left"                   
  )

variable_table
```


#### Background information

```{r provision-counts, echo=FALSE, results='hide'}

n_provisioned <- data %>%
  filter(condition != "wild") %>%
  nrow()

n_provisioned_not_excluded <- data %>%
  filter(condition != "wild", is.na(exclude_entry) | exclude_entry == "") %>%
  nrow()

n_provisioned_species <- data %>%
  filter(condition != "wild") %>%
  summarise(n_species = n_distinct(species_10ktrees)) %>%
  pull(n_species)

n_provisioned_s <- as.character(n_provisioned)
n_provisioned_not_excluded_s <- as.character(n_provisioned_not_excluded)
n_provisioned_species_s <- as.character(n_provisioned_species)

```

Background information variables are listed in Table `r tbl_context_num`. For each study, we provide information about the location and name of the field site, whether the animals were provisioned or not (n=`r n_provisioned_s` studies on `r n_provisioned_species_s` species reported that the animals were provisioned or commensal in some way), the years in which paternity data were collected, what sample matrix or matrices the DNA was obtained from, the number of loci that were evaluated, the paternity confidence levels that were used, and whether the analyses were primarily longitudinal or cross-sectional. As described above, we also noted whether there was known data overlap with other studies, and if so, which of the studies was the largest/newest. We did not encounter any cases where the newest publication did not also contain the largest number of paternities.

```{r study-context-table, echo=FALSE}

# Variables to include in Table 2

variables_of_interest_2 <- c(
  "country",
  "field_site",
  "condition",
  "year_first",
  "year_last",
  "n_yrs_sampled",
  "sample_matrix_1",
  "loci_lo",
  "loci_hi",
  "paternity_conf_strict",
  "paternity_conf_relax",
  "long_or_cross",
  "study_overlap",
  "study_overlap_name",
  "newest_biggest"
)

# Descriptions for Table 2 variables
descriptions_2 <- c(
  country = "Country where the study was conducted",
  field_site = "Name of the field site",
  condition = "Whether the population was provisioned or wild-feeding",
  year_first = "Earliest year included in study",
  year_last = "Latest year included in study",
  n_yrs_sampled = "Number of years sampled",
  `sample_matrix_1` = "Type of matrix used (e.g., feces, tissue)",
  loci_lo = "Minimum number of loci used",
  loci_hi = "Maximum number of loci used",
  paternity_conf_strict = "Paternity confidence level (strict)",
  paternity_conf_relax = "Paternity confidence level (relaxed)",
  long_or_cross = "Whether analysis was primarily cross-sectional or longitudinal",
  study_overlap = "Do paternity data in this study overlap with another study?",
  study_overlap_name = "If there is overlap with another study, which one?",
  newest_biggest = "If there is overlap with another study, is this study the newest/biggest?"
)

# Build/display table
variable_table_2 <- tibble(
  `Variable name` = variables_of_interest_2,
  `Variable description` = descriptions_2[variables_of_interest_2]
) %>%
  gt() %>%
  tab_header(
    title = paste0("Table ", tbl_context_num, ". Study context and sample characteristics")
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size = px(13),
    table.align = "left"
  ) %>%
  tab_footnote(
    footnote = "Actual table contains sample_matrix_1, sample_matrix_2, etc. up to sample_matrix_6, because some studies include DNA from multiple sources (e.g. feces and tissue).",
    locations = cells_body(rows = `Variable name` == "sample_matrix_1")
  )

variable_table_2
```


#### Study subject characteristics

Study subject variables are listed in Table `r tbl_taxonomy_num`. We include the common and scientific species name provided by the authors of the study, as well as the genus, species, and where applicable subspecies name used by the 10kTrees website [@Arnoldetal2010] used to generate our phylogenetic distance matrix (necessary for answering our question about the degree to which phylogeny predicts paternity loss, and to control for phylogenetic relatedness when answering other questions). In most cases these names were identical. Occasionally they were different due to updates in classification (e.g., 10kTrees uses the older, monogeneric _Callicebus_ classification for titi monkeys, while one of the studies we included uses _Plecturocebus_ [@Dolotovskayaetal2020], because 10kTrees did not contain an exact match (e.g., it did not have _Brachyteles hypoxanthus_, so we used _B. arachnoides_), or because the study did not use a subspecies name while 10kTrees required it (e.g., few published studies distinguish between _Pan troglodytes_ subspecies, while 10kTrees requires specification of either _P. t. versus_ or _P. t. schweinfurthii_). 

We also recorded the names of the social groups whenever they were provided; in cases where numbers from multiple groups were aggregated, we recorded the group name as "multiple_" followed by the name of the field site (e.g. multiple_Amboseli). We also classified groups based on the number of adult residents of each sex living together as a cohesive unit, rather than the presumed mating system for the taxa. Group composition was categorized as follows: MF = one resident (adult) male and one resident (adult) female, which we refer to as pair-living; MFF = one resident male and two or more resident females; MMF = two or more resident males and one resident female; MMFF = two or more resident adults of each sex. These abbreviations are used in tables and figures throughout our analyses. In some cases, group composition changed across time (e.g. from MF to MFF), so we tabulated data separately for each group composition type whenever possible. 

```{r table-3-taxonomy, echo=FALSE}

# Variables for Table 3

variables_of_interest_3 <- c(
  "suborder",
  "infraorder",
  "parvorder",
  "family",
  "common_name",
  "sci_name_inpaper",
  "genus_10ktrees_str",
  "species_10ktrees_str",
  "subspecies_10ktrees_str",
  "genus_species_subspecies",
  "genus_10ktrees",
  "species_10ktrees",
  "subspecies_10ktrees",
  "social_grp",
  "grp_composition"
)

# Descriptions
descriptions_3 <- c(
  suborder = "Is the species a Haplorrhine or Strepsirrhine?",
  infraorder = "What infraorder does the species belong to?",
  parvorder = "What parvorder does the species belong to (if applicable)?",
  family = "What family does the species belong to?",
  common_name = "Species common name",
  sci_name_inpaper = "Scientific name authors used in paper (string variable)",
  genus_10ktrees_str = "Genus name used in the phylogenetic matrix (string variable)",
  species_10ktrees_str = "Species name used in the phylogenetic matrix (string variable)",
  subspecies_10ktrees_str = "Subspecies name used in the phylogenetic matrix (string variable)",
  genus_species_subspecies = "Species classification from 10kTrees (factor variable)",
  genus_10ktrees = "Genus name from 10kTrees (encoded)",
  species_10ktrees = "Species name from 10kTrees (encoded)",
  subspecies_10ktrees = "Subspecies name from 10kTrees (encoded)",
  social_grp = "Name of the social group (when known)",
  grp_composition = "Group composition (MF, MFF, MMF, MMFF)"
)

# Build/display the table
variable_table_3 <- tibble(
  `Variable name` = variables_of_interest_3,
  `Variable description` = descriptions_3[variables_of_interest_3]
) %>%
  gt() %>%
  tab_header(
    title = paste0("Table ", tbl_taxonomy_num, ". Study subject characteristics")
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size = px(13),
    table.align = "left"
  ) %>%
  tab_footnote(
    footnote = "When data from multiple groups are aggregated, the entry is recorded as multiple_name of field site (e.g., multiple_Amboseli).",
    locations = cells_body(rows = `Variable name` == "social_grp")
  )

variable_table_3
```


#### Information about paternity distribution

Paternity distribution measures are listed in Table `r tbl_paternity_num`. Authors use different metrics to report paternity results, but for the majority of studies it is possible to tabulate the total number of paternities tested along with (a) the number of paternities obtained by the resident male in groups that include only one male, or (b) the number of paternities obtained by the top-ranking (alpha) male in groups that include more than one male. We categorized (a) and (b) together as primary males. We also extracted the number sired by other group residents of 1) known rank and 2) unknown rank. Note that the second category includes studies where the rank(s) of the males in question were unknown due to lack of behavioral data, where they were not reported, or where the authors were unable to conclusively assign a single in-group male as the father. In addition, we extracted the number of individuals that were sampled but for whom paternity remained undetermined, number of extra-group paternities, number where whether the paternity was extra-group was ambiguous, number sired by the single most successful male in the study, and number sired by all other (i.e., not most successful) males combined. These categories are not necessarily mutually exclusive, so the summed totals across rows do not necessarily match the total number of individuals that were sampled. For example, a given paternity could appear in both the column that tabulates the number of individuals for whom paternity was undetermined and in the column that tabulates instances where extra-group paternity was ambiguous, because the authors identified two equally likely potential sires but did not have demographic data to determine whether these potential sires were residents of the group at the time of conception.

In addition to such count data, some studies report other metrics such as the mean rank of sires or the B-index of reproductive skew [@Nonacs2000]. While Nonac's B-index (NB) suffers from flaws (namely, increasing values in larger group sizes are a statistical artifact of the way the index is constructed [@KutsukakeNunn2006; @Rossetal2020]), it remains the most commonly reported skew index value in the primate literature. We did not recalculate NB or mean rank of sires if they were not provided by the authors, because the necessary data to make these calculations were often unavailable. In cases where multiple NB values were included in a single study (e.g. at multiple points in time), we recorded all available NB values, confidence intervals, and p-values. When possible, we also recorded whether or not the authors had genetic data for all of the resident males.

No single study had all of these metrics available, and some had only a small number of them. When we had meaningful doubt about our ability to accurately calculate or extract a desired metric, we emailed the first author of the study to request clarification. In some cases they did not respond, in which case we left out a number that in theory could have been calculated. If the authors of any of our included studies find mistakes or fillable gaps in the table, we welcome the chance to update our existing information and to add new information. For studies where so little information was extractable that they cannot be included in any of the present analyses, we still recorded them in the interest of maintaining a complete list of studies that have primate genetic paternity data.

```{r table-4-skewmetrics, echo=FALSE}

# Variables for Table 4

variables_of_interest_4 <- c(
  "n_paternities",
  "n_alphaprimary",
  "n_residents_knownrank",
  "n_residents_unknownrank",
  "n_tested_undetermined",
  "n_egp",
  "n_egp_ambiguous",
  "n_sired_mostsuccess",
  "n_sired_othermales",
  "mean_sire_rank",
  "meanrank_standard",
  "nonacs_b",
  "ciforbindex_low",
  "ciforbindex_high",
  "nonacsb_p"
)

# Descriptions for each variable
descriptions_4 <- c(
  n_paternities = "Number of paternities tested",
  n_alphaprimary = "Number of paternities attributed to the alpha or resident male",
  n_residents_knownrank = "Number of paternities attributed to resident males of known rank",
  n_residents_unknownrank = "Number of paternities attributed to resident males of unknown rank",
  n_tested_undetermined = "Number of paternities that were tested but specific sire was not determined",
  n_egp = "Number of established extra-group paternities (EGPs)",
  n_egp_ambiguous = "Number of paternities that cannot be confidently assigned as either EGPs or resident sirings",
  n_sired_mostsuccess = "Number of paternities attributed to the most successful male",
  n_sired_othermales = "Number of paternities attributed to all other (not most successful) males",
  mean_sire_rank = "Mean dominance rank of sires, when ranks are known",
  meanrank_standard = "Is mean rank of sires standardized (Y/N)?",
  nonacs_b = "Nonac's B index of reproductive skew",
  ciforbindex_low = "Lower CI for Nonac's B",
  ciforbindex_high = "Upper CI for Nonac's B",
  nonacsb_p = "p-value for Nonac's B"
)

# Build/display the table
variable_table_4 <- tibble(
  `Variable name` = variables_of_interest_4,
  `Variable description` = descriptions_4[variables_of_interest_4]
) %>%
  gt() %>%
  tab_header(
    title = paste0("Table ", tbl_paternity_num, ". Paternity distribution information")
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size = px(13),
    table.align = "left"
  ) %>%
  tab_footnote(
    footnote = "When a study contained Nonac's B-index values from >1 point in time, these are recorded as nonacs_b_t2 and nonacs_b_t3. The same applies to the ciforbindex_low and _high entries (which contain the low and high confidence intervals associated with a B-index at t, t2, t3), and to the nonacsb_p entries (which contain the p values associated with the B-indexes).",
    locations = cells_body(rows = `Variable name` == "nonacs_b")
  )

variable_table_4
```

#### Information about the species

Species-level information is listed in Table `r tbl_traits_num`. This was almost always obtained from sources other than the genetic paternity studies in our table; much of it came from @KamilarCooper2013 and @Lupoldetal2019. We recorded whether they were seasonal breeders, whether they live in a multilevel society, morphological data relevant to dimorphism and sexually selected traits, information about species-typical adult group member composition, and the adult sex ratio. Not all measures were available for all species. Many of the available measures rely on older data or small sample sizes. If experts on a given species are able to provide more up-to-date species-level information, we welcome the opportunity to improve our table.

```{r table-5-speciestraits, echo=FALSE}

# Variables for Table 5

variables_of_interest_5 <- c(
  "seasonality_5c*",
  "seasonality_5c_2*",
  "seasonality_3c*",
  "multilevel",
  "philopatry_pattern",
  "pairbond_type",
  "gestation_length",
  "testes_mass",
  "litter_size",
  "mean_rainfall_mm",
  "mean_male_mass",
  "mean_female_mass",
  "mass_dimorphism",
  "canine_dimorphism",
  "mean_grp_size",
  "mean_num_males",
  "mean_num_females",
  "adult_sex_ratio"
)

# Corresponding descriptions
descriptions_5 <- c(
  "seasonality_5c*" = "Degree of birth seasonality in natural habitat (1=most, 5=none)",
  "seasonality_5c_2*" = "Degree of birth seasonality in natural habitat (1=most, 5=none)",
  "seasonality_3c*" = "Degree of birth seasonality in natural habitat (1=most, 3=none)",
  "multilevel" = "Does species live in multilevel society?",
  "philopatry_pattern" = "Primarily male or female philopatric?",
  "pairbond_type" = "In MF groups, are pairs dispersed or cohesive?",
  "gestation_length" = "Gestation length (days)",
  "testes_mass" = "Combined weight of testicles (g)",
  "litter_size" = "Mean litter size",
  "mean_rainfall_mm" = "Mean precipitation (mm)",
  "mean_male_mass" = "Average male weight (g)",
  "mean_female_mass" = "Average female weight (g)",
  "mass_dimorphism" = "Weight dimorphism value",
  "canine_dimorphism" = "Canine dimorphism value (canine length)",
  "mean_grp_size" = "Average size of social group",
  "mean_num_males" = "Average number of males who live in a social group",
  "mean_num_females" = "Average number of females who live in a social group",
  "adult_sex_ratio" = "Ratio of adult females to adult males"
)

# Build/display table
variable_table_5 <- tibble(
  `Variable name` = variables_of_interest_5,
  `Variable description` = descriptions_5[variables_of_interest_5]
) %>%
  gt() %>%
  tab_header(
    title = paste0("Table ", tbl_traits_num, ". Ecological, morphological, and life history traits")
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size = px(13),
    table.align = "left"
  ) %>%
  tab_footnote(
    footnote = "Birth seasonality classifications were obtained from Heldstab 2021. Heldstab includes three different classification schemes; our dataset includes all three.",
    locations = cells_body(rows = `Variable name` %in% c("seasonality_5c*", "seasonality_5c_2*", "seasonality_3c*"))
  )

variable_table_5
```


### Frequency of extra-group paternities (EGPs)

Many studies report the number of paternities obtained by males from outside the group (extra-group paternities, EGP). In some cases, paternity was assigned to specific extra-group males who had been sampled. In other cases, all resident males were sampled and could be excluded from paternity, and we assigned these paternities to extra-group males. If not all potential sires from within the group were sampled, paternities that could not be assigned to known male residents were categorized as ambiguous (variable n_egp_ambiguous). 

### Proportion of conceptions by primary male and extra-group males 

For cases in which all paternities could be assigned to primary males, other resident males, or extra-group males, the exact proportions of paternities by the primary male, other residents, and extra-group males were calculated by dividing the number of paternities in each category by the sum of paternities by primary males, other residents, and extra-group males.

In cases in which some paternities were categorized as residents of unknown rank, we calculated minimum and maximum estimates of the proportion of paternities obtained by primary males. To calculate the minimum estimate, we assumed that none of paternities attributable to residents of unknown rank were sired by primary males. We divided this number of paternities by the sum of all paternities (i.e., infants sired by the primary male, other residents, EGPs, and cases where it was ambiguous whether it was an EGP or in-group siring). To calculate the estimate of the maximum proportion sired by primary males, we assumed that all paternities in this category were sired by primary males, and divided by the same sum to obtain the proportion sired by primary males. The actual figure lies somewhere in between these two values. 

Similarly, in cases in which some paternities were categorized as “ambiguous”, we calculated minimum and maximum estimates of the proportion of paternities by extra-group males. To calculate the minimum estimate, we assumed that all paternities in this category were sired by resident males, and divided the number of paternities by extra-group males by the sum of total paternities. To calculate the maximum estimate, we assumed that all paternities in this category were sired by extra-group males, and divided by the same sum. The actual figure lies somewhere in between these two values.  

### Theoretical model

To clarify what data would ideally be needed to test the relationship between competition and paternity, and to make explicit the assumptions that underlie our empirical approach, we developed a simple formal theoretical framework. We assume that the proportion of paternities ($p$) that a primary male in a social group will obtain is, at its most basic, a function of two things: first, the number of competitors that he has ($n$), and second, the degree to which the mating opportunities in his group are monopolizable ($m$) [e.g., @KutsukakeNunn2006]. In other words:

$$
p=f(n,m)
$$

We will assume that as $n$ increases, $p$ should decrease (because more competitors likely equals more paternities lost), and that as $m$ increases, $p$ should increase (because greater monopolizability should mean more paternities gained). A simple model of this relationship could be:

$$
p=\frac{m}{m+n}
$$
Figure `r fig_simulation` contains a visualization of these dynamics. Note that this is in most respects just a recapitulation of Altmann’s priority-of-access model [@Altmann1962], but it 1) focuses specifically on the paternity share of one animal (i.e., the primary male), 2) replaces the number of estrus females with fertile female:male ratio, and 3) makes no assumptions about the competitive abilities of individual males.


```{r monopolizability-simulation-plot, fig.width=10, fig.height=4, fig.align="left", echo=FALSE, fig.cap=fig_simulation_cap}

# Create 105 observations
n_obs <- 105
N <- rep(0:20, times = 5)                     # 21 values (0 to 20) for each ratio
group <- rep(1:5, each = 21)                  # Assign group IDs
ratio <- recode(group, `1` = 5, `2` = 2, `3` = 1, `4` = 0.5, `5` = 0.2)  # Fertile F:M ratios
M_simulated <- 1 / ratio                      # Higher M = more monopolizability
P_simulated <- M_simulated / (M_simulated + N)  # Proportion paternities

# Combine into a data frame
sim_data <- data.frame(N, group, ratio, M_simulated, P_simulated)

# Define color mapping
color_map <- c("5:1 (Low monopolizability)" = "#56B4E9",
               "2:1" = "#009E73",
               "1:1" = "#E69F00",
               "1:2" = "#D55E00",
               "1:5 (High monopolizability)" = "#CC79A7")

# Prepare labeled data
sim_data <- sim_data %>%
  mutate(
    ratio_factor = factor(
      case_when(
        ratio == 5   ~ "5:1 (Low monopolizability)",
        ratio == 2   ~ "2:1",
        ratio == 1   ~ "1:1",
        ratio == 0.5 ~ "1:2",
        ratio == 0.2 ~ "1:5 (High monopolizability)"
      ),
      levels = names(color_map)
    )
  )

# Plot
ggplot(sim_data, aes(x = N, y = P_simulated, group = ratio_factor)) +
  geom_line(aes(color = ratio_factor), linewidth = 1.2) +
  scale_color_manual(values = color_map, name = "Monopolizability ratio") +
  labs(
    title = "Effect of competitors and monopolizability on paternity",
    x = "Number of competitors (n)",
    y = "Proportion of paternities (p)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    plot.title = element_text(size = 13, face = "plain"),
    panel.grid.minor = element_blank()
  )
```

Since competition can come from either males within the same social group or males from outside of it, $n$ can further be broken out as follows:

$$
p=\frac{m}{m + \alpha n_{\text{in}} + \beta n_{\text{out}}}
$$

where $n_{\text{in}}$ is the number of within-social group competitors a male faces, $n_{\text{out}}$ is the number of extra-group competitors he faces, and $\alpha$ and $\beta$ are weighting parameters that indicate the relative strength of intra- and extra-group competitors (e.g., in mountain gorillas, extra-group competitors seem to be harder to drive off, perhaps because they are more motivated to fight for mating opportunities [@RobbinsSawyer2007]).

The value of having the theoretical framework is that it makes it clear precisely what data would be needed for a rigorous test. Specifically, we would need: (1) conception-level data on social conditions at the time each offspring was sired, rather than paternity proportions aggregated over long time periods; (2) a continuous measure of the number of competitors, including some way to separately quantify within-group ($n_{\text{in}}$) and extra-group ($n_{\text{out}}$) competition; and (3) a direct measure of monopolizability at the time of conception, rather than a species-level proxy. 

Few papers that contain paternity data report this information. Due to primates’ slow life histories and the realities of sample collection and analysis, paternity determination usually does not happen until months or even years after offspring are born. The social conditions of the father at the time of conception of a given offspring are frequently omitted from studies, sometimes because they are not available, and sometimes because they are not of interest. Small sample sizes mean that often data are aggregated over many years. However, unless group composition and number of mating opportunities are extremely stable, statistics like the proportion of infants sired by primary male(s) across long time periods hides the potentially considerable amount of variation in social conditions that may have occurred at the point of the conception, which is the biologically relevant time window.

Given these constraints, our empirical analyses depart from the theoretical model in several ways. We operationalized the number of competitors as a binary indicator of group composition (single-male vs. multi-male), which collapses the (ideally) continuous $n_{\text{in}}$ term to a categorical variable and leaves $n_{\text{out}}$ unmeasured. For monopolizability, we used a species-level index of reproductive seasonality [@Heldstabetal2021]. This assumes that $m$ is a reasonable proxy for the temporal distribution of fertile periods, a potentially questionable assumption [see, e.g., @Ostneretal2008; @GogartenKoenig2013] but also ignores others potentially important factors such as female spatial distribution. 

While these metrics are much less precise than individual conception-level information would be, they are reasonable proxies given the constraints of  currently available data. The theoretical model helps make these compromises and their implications transparent. More details on our empirical implementation are provided below in the data analysis section. 

### Data analysis

All analyses were run in R 4.3.2 and all models were fit using brms [@Burkner2017], a flexible software package for fitting Bayesian multilevel models with a variety of response distributions, link functions, and prespecified correlation structures. These features make brms well-suited to our phylogenetic analyses of paternity patterns. We first constructed a phylogenetic covariance matrix using the consensus primate tree available from 10kTrees [@Arnoldetal2010] that is included in all models below as a random effect with a covariance structure corresponding to evolutionary distances between species (see [@Burkner2025] for a vignette of these models in brms). 

The proportion of paternity for a given class of males (i.e., primary, resident, extra-group) can span the entire range from 0 to 1. While Beta regression is a popular option to model proportions as response variables, it cannot accommodate proportions at either extreme–i.e., exactly 0 or exactly 1, a situation which is common in our data set. While there are modifications that can accommodate values of exactly 0 and/or 1 (e.g., zero-one inflated Beta [@LiuKong2015]), an underlying assumption of such models is that the processes that produces 0s, 1s, or both are different in some way than the processes that produce the rest of the distribution. We do not think that this is a biologically defensible stance for our analysis. The fundamental processes that produce 0s and 1s are likely the same as the ones that produce less extreme values. Furthermore, _true_ 0s and 1s are probably unlikely to occur in the real world. It seems more  likely that our 0 and 1 values are the result of small sample sizes and the potential for biased sampling of certain paternities, than that there are species in the real world where primary males either obtain every possible paternity or never obtain any of them. Thus, in order to use a standard Beta regression model, we replaced observed proportions of 1 with 0.999, and observed proportions of 0 with 0.001.  

For our first analysis examining the influence of phylogeny alone (Question 1 in the results), we specified our models predicting our various operationalizations of paternity proportions as:

$$
\text{logit}(\mu_i) = \tau_{\text{phylo}[i]} + \tau_{\text{species}[i]}
$$
where the observation-level paternity proportions ($\mu_i$) are predicted by random effects ($\tau$) for phylogeny and species. The $\tau_{\text{phylo}[i]}$ term captures variation due to phylogenetic relatedness between species, and its variance quantifies the amount of interspecies variation that is attributable to phylogeny. The $\tau_{\text{species}}$ term estimates interspecific variance not explained by phylogenetic relatedness. From these estimates, we calculated the phylogenetic signal ($\lambda$) by dividing the total variance from phylogeny  ($\sigma^2_{\text{phylo}}$) by the sum of the variance attributable to phylogeny plus the variance attributable to species-specific variation that is not explained by phylogeny:

$$
\lambda = \frac{\sigma^2_{\text{phylo}}}{\sigma^2_{\text{phylo}} + \sigma^2_{\text{nonphylo}}}
$$

Thus, $\lambda$ is the proportion of interspecies variance that is phylogenetically structured. Another way to state this is that it answers the question “of the variation in primary male paternity share between species, how much of it is explained by phylogeny?” This is analogous but not identical to Pagel's $\lambda$ [@Pagel1999], which is estimated by directly transforming the phylogenetic tree to find the best fit to the data and incorporates total residual variance. Unlike Pagel's $\lambda$, our $\lambda$ is a ratio of the variance components that are derived from the model's random effects, and excludes observation-level variance from the Beta distribution. We chose this approach because it directly addresses the between-species question and benefits from the observation weights and Beta likelihood included in the model.

Our second question is about the role that group composition and seasonality play in primary males' paternity share (Question 2 in the results). To evaluate their effects, first, we conducted a basic analysis that compared primary male paternity share values in multi-male, single-male, and pair-living groups; we present the raw data but also model-derived estimates that account for phylogeny and sample size (Analysis 1 in the results). Second, to determine how the combined influences of competition and seasonality affect paternity share (Analysis 2 in the results), we turned to our theoretical model described above ($p=m/(m+n)$). This model assumes a fixed functional relationship among the variables--specifically, that paternity share is determined by the ratio of monoplizability to competitive pressure exerted by other males. This is a useful conceptual foundation but may be too restrictive to capture the range of empirical variation observed across species. To allow for more flexibility, including potential non-linear effects of competitors and monopolizability, we constructed a model that contained main effects for both $n$ and $m$ plus an interaction between the two, as well as a squared term for $m$, while retaining the same random effects structure described above:

$$
\text{logit}(\mu_i) = \alpha + \beta_1 n_i + \beta_2 m_i + \beta_3 m_i^2 + \beta_4 (n_i \cdot m_i) + \tau_{\text{phylo}[i]} + \tau_{\text{species}[i]}
$$
While ideally we would also include a squared term for $n$, because we were only able to operationalize competitors as a categorical variable where 0=no in-group competitors (i.e., males living in single-male groups) and 1=one or more in-group competitors (i.e., males living in multi-male groups), a squared term would be entirely redundant with the main effect term. Monpolizability was operationalized as a reproductive seasonality variable obtained from @Heldstabetal2021. For this, lower values indicate more seasonality (and thus lower monpolizability) and higher values equal less seasonality (and thus higher monpolizability). To interpret the results we plotted the marginal effects and computed the partial derivative ($\partial$) of $logit(\mu_i)$ with respect to monopolizability ($\frac{\partial \, \text{logit}(\mu_i)}{\partial m}$), evaluated at the mean of $m$. For reporting purposes we transformed it to the response scale via the chain rule, which provides the estimated change in paternity share per unit increase in monopolizability. For the binary competitor variable $n$, we interpret the contrast in predicted values between single-male and multi-male groups.

To answer our third question, about who alpha males in multi-male groups lose paternities to (Question 3 in the results), we fit a binomial generalized linear mixed model (GLMM). Here, the outcome variable is the number of EGPs for a given observation, treated as the number of "successes" out of total count of paternities not attributed to the primary male (i.e., EGPs plus sirings by non-alpha resident males). Formally, we specified this as:

$$
\text{EGP}_i \sim \operatorname{Binomial}(n_i, \pi_i) \\
\text{logit}(\pi_i) = \alpha + \tau_{\text{phylo}[i]} + \tau_{\text{species}[i]}
$$

where $\text{EGP}_i$ is the observed number of EGPs for observation $i$, $n_i$ is the total number of non-alpha paternities for that observation, and $\pi_i$ is the probability that a non-alpha paternity is attributable to an extra-group male. For this model we kept our zero-loss observations as zeros rather than rounding to 0.001, because in this case we were modeling counts (i.e., "0 trials out of X trials") rather than a continuous distribution of probabilities. Again, zeros likely do not represent a fundamentally different biological sampling process as non-zeros, so we elected not to use (e.g.) a hurdle or zero-inflated model.  

All Beta regression models were weighted by the square root of the total number of paternities in a given study, to assign greater weight to larger studies that had more precise estimates of paternity distribution proportions. For the binomial model (Question 3), no additional weighting was applied because the binomial likelihood already accounts for sample size through the number of trials. We set a Gaussian (normal) prior on the intercept, centered at 0 (SD=1) on the logit scale. This means we set a weak prior centered on primary males obtaining 50% of paternities, while allowing for significant uncertainty and a wide range of plausible values. For the standard deviation of the random effects and the distribution precision parameter ($\phi$), we used Student’s t-distribution priors (df=3, center=0, scale=2.5). These are weakly informative, to allow for the possibility of substantial species-level variation and data dispersion, both of which we observed in our initial data exploration. We used the same strategy for the binomial model, except that the precision parameter was excluded.   

### Living table and results

Across scientific disciplines, researchers have recognized the value of building "living" systematic reviews and/or databases. While there are numerous tools, techniques, and strategies for providing these resources, all of these approaches share a common goal of creating scientific products that facilitate mechanisms for update [see e.g., @Elliottetal2017; @Mochaetal2024]. In deciding how to structure the database and the results derived from it in a continually updateable format, we drew inspiration from the Cochrane Guidance for Living Systematic Reviews [@Brookeretal2019]. Below are the core features of our living resources:

1) The results reported below in this static manuscript are based on the current version of the table as of March 2026. 

2) Twice a year, we will perform a new search for relevant data, using a strategy identical to the one described above in Data compilation, and we will accordingly update the table with any new data published since the last search.

3) If new data are identified, the relevant features are extracted, screened, and aggregated according to the methods described above.

4) We provide the most current version of the paternity dataset on an open-access Github repository (https://github.com/slrosen/primate-paternity-metanalysis). We document all updates made to the underlying dataset in a README file. Additionally, in the dataset we include a column listing the date each row was added or updated (variable dateofentry). 

5) The most current version of our results, as well as any previous versions, are also posted in the same open-access Github repository.

## Results

**Description of the data set**

```{r basic-data-description, echo=FALSE, results='hide'}

# All entries in the dataset (including excluded ones)
n_total_entries <- nrow(data)

# Entries excluded from analyses because they are redundant or irrelevant
n_excluded <- sum(grepl("^Yes", data$exclude_entry, ignore.case = FALSE) & 
        !grepl("no total n, see notes", data$exclude_entry, ignore.case = TRUE), na.rm = TRUE)

# Entries included in analyses (exclude_entry is NA or empty)
data_clean <- data %>%
  filter(is.na(exclude_entry) | exclude_entry == "")
n_analysis_entries <- nrow(data_clean)

# Species, genera, and families across all entries (including excluded)
n_species_total <- n_distinct(data$species_10ktrees) # use this version so it doesn't count subspecies
n_families <- n_distinct(data$family)
n_genera <- n_distinct(data$genus_10ktrees)

# Total paternities, excluding entries marked as not newest/biggest or otherwise redundant 
data_nonoverlapping <- data %>%
  filter(!(exclude_entry %in% c("Yes (not newest/biggest)", "Yes (redundant, see notes)")))
total_paternities <- sum(data_nonoverlapping$n_paternities, na.rm = TRUE)

# Paternities and species actually included in the present analyses (remove everything in the exclusion table)
data_analysis <- data %>%
  filter(is.na(exclude_entry) | exclude_entry == "") %>%
  filter(!is.na(n_alphaprimary))  
analysis_paternities <- sum(data_analysis$n_paternities, na.rm = TRUE)
analysis_species <- n_distinct(data_analysis$species_10ktrees_str) # note this gets used in inline text later in place of the n from a model b/c the model is counting at subspecies level

# Entries missing critical info (primary male siring counts, total n)
n_missing_primary <- sum(is.na(data_clean$n_alphaprimary))
n_missing_no_total_n <- sum(grepl("no total n", data$exclude_entry, ignore.case = TRUE), na.rm = TRUE)
n_missinginfo <- n_missing_primary + n_missing_no_total_n
n_modeled <- n_analysis_entries - n_missing_primary

species_counts <- data_analysis %>%
  count(genus_species_subspecies, name = "n") # use this version so it matches tree data (needs subspecies)

# species in tree but not in the data
to_drop <- setdiff(phylo_tree$tip.label,
                   species_counts$genus_species_subspecies)

phylotree_subset <- drop.tip(phylo_tree, to_drop)

# turn tree into a ggtree object
gtre <- ggtree(phylotree_subset)

# add sample size info
gtre$data <- gtre$data %>%
  left_join(species_counts,
            by = c("label" = "genus_species_subspecies"))

# unique species-family pairs from the full dataset

species_family <- data_analysis %>%
  distinct(genus_species_subspecies, family)

# add family info to species_counts, and keep only taxa in the plotted tree

species_counts_trimmed <- species_counts %>%
  left_join(species_family, by = "genus_species_subspecies") %>%
  filter(genus_species_subspecies %in% phylotree_subset$tip.label)

# tbl_exclusion number is pre-computed in setup chunk

# ---- Data transformations used in multiple downstream analyses ----

# Add row IDs to track observations across analyses
data_clean <- data_clean %>%
  mutate(row_id = row_number())

# Subset data to multi-male and single-male groups
data_subset_mm <- data_clean %>%
  filter(grp_composition %in% c("MMF", "MMFF"))  

data_subset_sm <- data_clean %>%
  filter(grp_composition %in% c("MF", "MFF"))

# Recode group composition: combine MMF+MMFF into MM (used in Q2 Analysis 1 and Q3)
data_clean$grp_composition_recode <- as.character(data_clean$grp_composition)
data_clean$grp_composition_recode[data_clean$grp_composition_recode %in% c("MMF", "MMFF")] <- "MM"
data_clean$grp_composition_recode <- factor(data_clean$grp_composition_recode)
data_clean$grp_composition_recode <- relevel(data_clean$grp_composition_recode, ref = "MM")

# Binary recode: MM vs SM, with numeric dummy variable n (used in Q2 Analysis 2)
data_clean <- data_clean %>%
  mutate(
    grp_composition_recode_binary = case_when(
      grp_composition %in% c("MMF", "MMFF") ~ "MM",
      grp_composition %in% c("MF", "MFF")   ~ "SM",
      TRUE ~ as.character(grp_composition)
    ),
    grp_composition_recode_binary = factor(
      grp_composition_recode_binary, 
      levels = c("MM", "SM")
    ),
    n = if_else(grp_composition_recode_binary == "SM", 1, 0)
  )

# Code monopolizability variable (seasonality) and its quadratic (used in Q2 Analysis 2)
data_clean$m <- data_clean$seasonality_3c
data_clean$m2 <- data_clean$m^2

# Define fig_tree_cap here (rather than in setup) because it needs data-derived values
fig_tree_cap <- paste0("Figure ", fig_tree_num, ". Phylogenetic distribution and sample sizes of species in the database. The left panel shows the consensus phylogeny from 10kTrees for all species represented in the dataset. The right panel shows the number of observations (rows in the dataset) per species, colored by family. The sample size in this figure does not include n=", n_excluded + n_missinginfo, " observations that were excluded from our analyses for the reasons listed in Table ", tbl_exclusion_num, ". ")

```


Our database contains `r n_total_entries` entries from `r n_species_total` species representing `r n_genera` genera and `r n_families` families, encompassing a total of `r format(total_paternities, big.mark = ",")` non-overlapping paternities. Of these entries, `r n_excluded` were excluded up front either because of data redundancies or because they were irrelevant for our questions, while an additional `r n_missinginfo` lacked sufficient information to calculate primary male paternity share (see Table `r tbl_exclusion_num`). This left `r n_modeled` entries, `r analysis_species` species, and `r format(analysis_paternities, big.mark = ",")` paternities for the analyses presented below. The phylogenetic distribution of the species in our analysis and the number of observations per species are shown in Figure `r fig_tree`.

```{r exclusion-summary-table, echo=FALSE}

# Build the exclusion reasons from the exclude_entry column
exclusion_rows <- data %>%
  filter(grepl("^Yes", exclude_entry)) %>%
  count(exclude_entry) %>%
  mutate(Reason = case_when(
    grepl("not newest", exclude_entry, ignore.case = TRUE) ~ "Study not newest/biggest",
    grepl("redundant", exclude_entry, ignore.case = TRUE) ~ "Redundant entry",
    grepl("no total n", exclude_entry, ignore.case = TRUE) ~ "No total n",
    TRUE ~ "Group type/rank not applicable (genus *Pongo* only)"
  )) %>%
  select(Reason, n)

# Add row for missing primary male data
primary_row <- tibble(
  Reason = "Count of primary male sirings unavailable",
  n = n_missing_primary
)

exclusion_all <- bind_rows(exclusion_rows, primary_row) %>%
  mutate(Reason = factor(Reason, levels = c(
    "Study not newest/biggest",
    "Redundant entry",
    "Group type/rank not applicable (genus *Pongo* only)",
    "Count of primary male sirings unavailable",
    "No total n"
  ))) %>%
  arrange(Reason) %>%
  mutate(Reason = as.character(Reason))

exclusion_all %>%
  gt() %>%
  tab_header(
    title = paste0("Table ", tbl_exclusion_num, ". Reasons table entries were excluded from current analyses")
  ) %>%
  cols_label(
    n = "N"
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", align = "left"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_body(columns = everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size = px(13),
    table.align = "left"
  ) %>%
  tab_footnote(
    footnote = md("For 16 entries, all from a single study of *Hapalemur griseus* (Neivergelt et al. 2002), we were unable to conclusively determine the total number of paternities tested for any social group. Thus, we could not calculate proportion sired by primary males."),
    locations = cells_body(columns = Reason, rows = Reason == "No total n")
  ) %>%
  tab_footnote(
    footnote = md("For four entries, the information contained in the entry was an alternative breakdown of another entry in the table (e.g., data aggregated at the social group level that is aggregated in another entry at the study level)."),
    locations = cells_body(columns = Reason, rows = Reason == "Redundant entry")
  ) %>%
  tab_footnote(
    footnote = md("For 41 entries, we were unable to determine the number of offspring sired by alpha/primary males. This happened for a variety of reasons, most commonly because behavioral data were unavailable at the time the siring(s) occurred."),
    locations = cells_body(columns = Reason, rows = Reason == "Count of primary male sirings unavailable")
  ) %>%
  fmt_markdown(columns = Reason)

```

```{r tree-bar-figure, echo=FALSE, fig.cap=fig_tree_cap}

# tree plot

# Build unannotated tree
base_tree <- ggtree(phylotree_subset, layout = "rectangular")
tree_dat <- base_tree$data

# Join species info (n, family) onto the tip rows
tree_dat_annot <- tree_dat %>%
  left_join(
    species_counts_trimmed,
    by = c("label" = "genus_species_subspecies")
  )

# Species (tips) in vertical plotting order
tip_order <- tree_dat_annot %>%
  filter(isTip) %>%
  arrange(y) %>%
  pull(label)

# Families in *order of appearance down the tree* (unique in that order)
family_order <- tree_dat_annot %>%
  filter(isTip) %>%
  arrange(y) %>%
  pull(family) %>%
  unique() %>%
  rev() 


# Now build the final tree plot with family-colored tip dots
p_tree <- ggtree(phylotree_subset, layout = "rectangular") +
  # colored dots at tips, mapped to family
  # geom_tippoint(
  #  data = tree_dat_annot %>% filter(isTip),
  #  aes(x = x, y = y, color = family),
  #  size = 2,
  #  show.legend = FALSE  #hiding duplicate legend, will show the one from the bar plot
#  ) +
  # tip labels in black (readable), with our tuned spacing
  geom_tiplab(
    align      = FALSE,
    linetype   = "dotted",
    linesize = 0,
    size       = 2.8,
    lineheight = 0.8,
    offset     = 0.3,
    color      = "black"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0.02, 0.15))
  ) +
  coord_cartesian(clip = "off") +
  theme_tree2() +
  theme(
    plot.margin = margin(5, 50, 5, 5),
    axis.line   = element_blank(),
    axis.text   = element_blank(),
    axis.title  = element_blank(),
    axis.ticks  = element_blank()
  )
# p_tree is combined with p_bar below; not printed standalone

# Bar panel--start by making the order match

# Order of species along the y-axis of the tree
tip_order <- tree_dat_annot %>%
  filter(isTip) %>%
  arrange(y) %>%
  pull(label)

p_bar <- species_counts_trimmed %>%
  mutate(
    genus_species_subspecies = factor(
      genus_species_subspecies,
      levels = tip_order
    ),
    family = factor(
      family,
      levels = family_order
    )
  ) %>%
  ggplot(aes(
    x    = n,
    y    = genus_species_subspecies,
    fill = family
  )) +
  geom_col(width = 0.6) +
  labs(
    x = "Sample size",
    y = NULL,
    fill = "Family"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x        = element_text(size = 10),   
    axis.title.x       = element_text(size = 10, face = "bold"),  
    axis.text.y      = element_blank(),
    panel.grid       = element_blank(),
    plot.margin      = margin(5, 5, 5, 15),

    # put legend in the top-right inside the bar panel
    legend.position  = c(1, 1),            # (x, y) in npc coords of that panel
    legend.justification = c(1, 1),        # anchor the legend's top-right corner 

    # shrink legend appearance
    legend.title     = element_text(size = 10),
    legend.text      = element_text(size = 9),
    legend.key.size  = unit(0.7, "lines"),
    legend.spacing.y = unit(0.4, "lines"),
    legend.spacing.x = unit(0.4, "lines"),
    legend.key = element_rect(color = NA)
  ) +
  guides(
    fill = guide_legend(
      ncol = 1,                # vertical stack (one entry per row)
      byrow = FALSE
    )
  )

# Combine tree panel and bar panel
combined_plot <- p_tree + p_bar + 
  plot_layout(widths = c(2.2, 1)) +
  plot_annotation(title = "Phylogenetic distribution and sample sizes of species in the database")

# Save 
ggsave(
  "../output/plots/tree_sample_size_plot_final.pdf",
  combined_plot,
  width  = 8,
  height = 10   # ⬅ intermediate vertical height
)

combined_plot

```


**Question 1: To what degree does phylogeny predict the proportion of paternities obtained by primary males?**

```{r lambda-check, echo=FALSE, results='hide'}

#Checking to see if we get a lambda number for # males per group that is roughly the same as what Kamilar & Cooper reported with their larger data set. Lambda is sensitive to missing values and we filtered a lot of the species out that they used; do we still get the same thing?

# Load tree

tree_path <- file.path("..", "data", "phylogeny", "consensusTree_10kTrees_Primates_Version3.nex")
if (!file.exists(tree_path)) {
  stop("Tree file not found at: ", normalizePath(tree_path, winslash = "/"))
}
phylo_10k <- read.nexus(tree_path)

# Create subspecies-like labels in the data (append _x to species-only rows)
ss_df <- data_clean %>%
  filter(!is.na(mean_num_males), !is.na(genus_species_subspecies)) %>%
  mutate(
    lbl_raw = gsub(" ", "_", genus_species_subspecies),
    n_us = lengths(regmatches(lbl_raw, gregexpr("_", lbl_raw))),
    subspp_label = if_else(n_us == 1L, paste0(lbl_raw, "_x"), lbl_raw)
  ) %>%
  group_by(subspp_label) %>%
  summarise(mean_num_males = mean(mean_num_males, na.rm = TRUE), .groups = "drop")

# Create trait vector using these new labels
x <- setNames(ss_df$mean_num_males, ss_df$subspp_label)

# Rename species-only tips w/in the tree to look like subspecies (append _x)
tree_ss <- phylo_10k
is_species_only <- grepl("^[A-Z][a-z]+_[a-z]+$", tree_ss$tip.label)
tree_ss$tip.label[is_species_only] <- paste0(tree_ss$tip.label[is_species_only], "_x")

# Align 
common <- intersect(names(x), tree_ss$tip.label)
tree2  <- drop.tip(tree_ss, setdiff(tree_ss$tip.label, common))
x2     <- x[tree2$tip.label]

cat("Trait rows after collapse:", length(x), "\n")
cat("Tree tips after renaming:", length(tree_ss$tip.label), "\n")
cat("Matched (used) tips:", length(x2), "\n")   # <- ends up at 39 b/c 8 of the 47 species/subspecies used in the intercept_only_model do not have mean_num_males values

# Use this if I want to check who dropped on each side
# not_in_tree <- setdiff(names(x), tree_ss$tip.label)   # in data, missing in tree
# not_in_data <- setdiff(tree_ss$tip.label, names(x))   # in tree, missing in data
# head(not_in_tree); head(not_in_data)

# Sanity: names must line up 1:1 for phylosig
stopifnot(all.equal(names(x2), tree2$tip.label), sum(is.na(x2)) == 0)

# Pagel's lambda
res_lambda <- phylosig(tree2, x2, method = "lambda", test = TRUE)
res_lambda$lambda
 res_lambda$P     
 res_lambda       # full output

# Look at what taxa were included in the $\lambda$ analysis 
stopifnot(exists("x2"), exists("tree2"))  # must be defined by the pipeline 

included_tips <- names(x2)                 # labels used by phylosig()
length(included_tips)
included_tips[1:20]                        # just to verify these are the names where I added _x

# Drop subspecies/_x
species_stem <- function(lbl) {
  sub("^([A-Za-z]+_[a-z]+).*", "\\1", sub("_x$", "", lbl))
}

# Species-level list (unique Genus_species)
included_species <- sort(unique(species_stem(included_tips)))
length(included_species)
included_species[1:20]                     # verify I got rid of _x

# Table with details
included_tbl <- tibble::tibble(
  tip_label = included_tips,
  species   = species_stem(included_tips),
  is_species_x = grepl("_x$", included_tips),
  is_true_subspecies = grepl("^[A-Z][a-z]+_[a-z]+_[A-Za-z0-9]+$", included_tips),
  trait_value = unname(x2)
) %>%
  arrange(species, tip_label)

# View(included_tbl)  

# Species -> how many tips used (should be 1 for species-only; can be >1 if subspecies present)
by_species_counts <- included_tbl %>% count(species, name = "n_tips_used") %>% arrange(desc(n_tips_used))

# Optional code to save lists if needed for elsewhere
# writeLines(included_tips,   "lambda_included_tip_labels.txt")
# writeLines(included_species,"lambda_included_species.txt")
# readr::write_csv(included_tbl, "../output/tables/lambda_included_table.csv")
# readr::write_csv(by_species_counts, "../output/tables/lambda_included_by_species_counts.csv")

# Print summary
cat("Tips used (n):", length(included_tips), "\n")
cat("Unique species (n):", length(included_species), "\n")
head(by_species_counts, 10)


```


```{r max-primary-paternity-models, echo=FALSE, results='hide', fig.show='hide'}

#This model assumes maximum possible paternity for primary males

#Full sample (all group types)
if (file.exists("../output/model-objects/intercept_only_model.rds")) {
  intercept_only_model <- readRDS("../output/model-objects/intercept_only_model.rds")
} else {
  intercept_only_model <- brm(
    bf(proportion_primary_maximum_adj | weights(n_paternities^0.5) ~
         (1 | gr(genus_species_subspecies, cov = phylo_matrix)) +
         (1 | species_nonphylo)),
    family = Beta(),
    data = data_clean,
    data2 = list(phylo_matrix = phylo_matrix),
    prior = c(
      prior(normal(0, 1), class = "Intercept"),
      prior(student_t(3, 0, 2.5), class = "sd"),
      prior(student_t(3, 0, 2.5), class = "phi")
    ),
    chains = 4, iter = 10000, seed = 123,
    control = list(adapt_delta = 0.99)
  )
  saveRDS(intercept_only_model, "../output/model-objects/intercept_only_model.rds")
}

  summary(intercept_only_model)
  plot(intercept_only_model)
  pp_check(intercept_only_model, type="hist")
  pp_check(intercept_only_model, type="dens_overlay")

  n_obs_iom <- nobs(intercept_only_model)
  n_species_iom <- length(unique(
    sub("^([^_]+_[^_]+).*", "\\1", intercept_only_model$data$genus_species_subspecies) #keep from counting subspecies
  ))
  n_pats_iom <- sum(intercept_only_model$data$n_paternities, na.rm = TRUE)

  # Determine R2 
  bayes_R2_iom <- as.data.frame(bayes_R2(intercept_only_model))
  R2_est_iom <- round(bayes_R2_iom$Estimate[1], 2)
  R2_ci_lo_iom <- round(bayes_R2_iom$Q2.5[1], 2)
  R2_ci_hi_iom <- round(bayes_R2_iom$Q97.5[1], 2)

  # Extract posterior draws
  post_iom <- as_draws_df(intercept_only_model)
  
  # Extract and square standard deviations to get variances
  var_phylo_iom <- post_iom$`sd_genus_species_subspecies__Intercept`^2
  var_nonphylo_iom <- post_iom$`sd_species_nonphylo__Intercept`^2
  
  # Calculate lambda
  lambda_iom <- var_phylo_iom / (var_phylo_iom + var_nonphylo_iom)
  
  # Summarize lambda
  summary_lambda_iom <- quantile(lambda_iom, probs = c(0.025, 0.5, 0.975))
  summary_lambda_iom

```


```{r min-primary-paternity-models, echo=FALSE, results='hide', fig.show='hide'}

#This models assumes minimum possible paternity for primary males

#Full sample (all group types)
if (file.exists("../output/model-objects/intercept_only_model_min.rds")) {
  intercept_only_model_min <- readRDS("../output/model-objects/intercept_only_model_min.rds")
} else {
  intercept_only_model_min <- brm(
    bf(proportion_primary_minimum_adj | weights(n_paternities^0.5) ~
         (1 | gr(genus_species_subspecies, cov = phylo_matrix)) +
         (1 | species_nonphylo)),
    family = Beta(),
    data = data_clean,
    data2 = list(phylo_matrix = phylo_matrix),
    prior = c(
      prior(normal(0, 1), class = "Intercept"),
      prior(student_t(3, 0, 2.5), class = "sd"),
      prior(student_t(3, 0, 2.5), class = "phi")
    ),
    chains = 4, iter = 10000, seed = 123,
    control = list(adapt_delta = 0.99)
  )
  saveRDS(intercept_only_model_min, "../output/model-objects/intercept_only_model_min.rds")
}

  summary(intercept_only_model_min)
  plot(intercept_only_model_min)
  pp_check(intercept_only_model_min, type="hist")
  pp_check(intercept_only_model_min, type="dens_overlay")
  
  n_obs_iom_min <- nobs(intercept_only_model_min)
  n_species_iom_min <- length(unique(
    sub("^([^_]+_[^_]+).*", "\\1", intercept_only_model$data$genus_species_subspecies)
  ))
  n_pats_iom_min <- sum(intercept_only_model_min$data$n_paternities, na.rm = TRUE)
  
  # Determine R2 
  bayes_R2_iom_min <- as.data.frame(bayes_R2(intercept_only_model_min))
  R2_est_iom_min <- round(bayes_R2_iom_min$Estimate[1], 2)
  R2_ci_lo_iom_min <- round(bayes_R2_iom_min$Q2.5[1], 2)
  R2_ci_hi_iom_min <- round(bayes_R2_iom_min$Q97.5[1], 2)

  # Extract posterior draws
  post_iom_min <- as_draws_df(intercept_only_model_min)
  
  # Extract and square standard deviations to get variances
  var_phylo_iom_min <- post_iom_min$`sd_genus_species_subspecies__Intercept`^2
  var_nonphylo_iom_min <- post_iom_min$`sd_species_nonphylo__Intercept`^2
  
  # Calculate lambda
  lambda_iom_min <- var_phylo_iom_min / (var_phylo_iom_min + var_nonphylo_iom_min)
  
  # Summarize lambda
  summary_lambda_iom_min <- quantile(lambda_iom_min, probs = c(0.025, 0.5, 0.975))
  summary_lambda_iom_min

```
Species-level random effects, which includes both phylogenetic relatedness and other species-level differences, explained roughly 35-40% of the variation in the proportion of paternities obtained by primary males. The posterior mean $R^2$ values of intercept-only models for all group types when all uncertain paternities were assigned to the primary male (`r format(R2_est_iom, nsmall=2)`, 95% CI: `r format(R2_ci_lo_iom, nsmall=2)` - `r format(R2_ci_hi_iom, nsmall=2)`, n=`r n_obs_iom` observations, `r analysis_species` species, and `r format(n_pats_iom, big.mark = ",")` total paternities) was very similar to the $R^2$ value when none of the uncertain paternities were assigned to the primary male (`r format(R2_est_iom_min, nsmall=2)`, 95% CI: `r format(R2_ci_lo_iom_min, nsmall=2)` - `r format(R2_ci_hi_iom_min, nsmall=2)`).  

We next examined how this explained variation was apportioned between phylogenetic relatedness and other, non-phylogenetic species-level factors. If we assume that primary males sired all infants of uncertain paternity, a substantial proportion of the variation in primary male paternity share was explained by phylogenetic relatedness in the full sample ($\lambda$=`r format(round(summary_lambda_iom[2], 2), nsmall = 2)`, 95% CI: `r format(round(summary_lambda_iom[1], 2), nsmall = 2)` - `r format(round(summary_lambda_iom[3], 2), nsmall = 2)`). Assigning all uncertain paternities to other, non-primary males decreased $\lambda$ and increased 95% CIs ($\lambda$=`r format(round(summary_lambda_iom_min[2], 2), nsmall = 2)`, 95% CI: `r format(round(summary_lambda_iom_min[1], 2), nsmall = 2)` - `r format(round(summary_lambda_iom_min[3], 2), nsmall = 2)`). 

**Question 2: How do group composition and reproductive seasonality influence primary males' paternity share?**

```{r summary-stats-paternity-loss, echo=FALSE}

#Calculate mean, SD aggregated across group types

mean_val_primary <- mean(data_clean$proportion_primary_maximum_adj, na.rm = TRUE)
sd_val_primary <- sd(data_clean$proportion_primary_maximum_adj, na.rm = TRUE)
n_val_primary <- sum(!is.na(data_clean$proportion_primary_maximum_adj))
n_pats_primary <- sum(data_clean$n_paternities[!is.na(data_clean$proportion_primary_maximum_adj)], na.rm = TRUE)
n_species_primary <- n_distinct(data_clean$genus_species_subspecies[!is.na(data_clean$proportion_primary_maximum_adj)])

mean_val_primary_min <- mean(data_clean$proportion_primary_minimum_adj, na.rm = TRUE)
sd_val_primary_min <- sd(data_clean$proportion_primary_minimum_adj, na.rm = TRUE)
n_val_primary_min <- sum(!is.na(data_clean$proportion_primary_minimum_adj))
n_pats_primary_min <- sum(data_clean$n_paternities[!is.na(data_clean$proportion_primary_minimum_adj)], na.rm = TRUE)
n_species_primary_min <- n_distinct(data_clean$genus_species_subspecies[!is.na(data_clean$proportion_primary_minimum_adj)])

# Formatted versions for inline text
mean_val_primary_s <- sprintf("%.0f", mean_val_primary * 100)
sd_val_primary_s <- sprintf("%.0f", sd_val_primary * 100)
mean_val_primary_min_s <- sprintf("%.0f", mean_val_primary_min * 100)
sd_val_primary_min_s <- sprintf("%.0f", sd_val_primary_min * 100)

```


```{r table-paternity-loss, echo=FALSE, results='hide'}

#This puts the summary-stats-paternity-loss info into table format. Not going to display until further below.

# keep the NA filters identical everywhere
valid_rows <- function(df) {
  !is.na(df$proportion_primary_minimum_adj) & !is.na(df$proportion_primary_maximum_adj)
}

# 1) Main table by grp_composition 
main_tbl <- data_clean %>%
  group_by(grp_composition) %>%
  summarise(
    n         = sum(valid_rows(cur_data_all())),
    n_species = n_distinct(species_10ktrees[valid_rows(cur_data_all())]),
    n_pats    = sum(n_paternities[valid_rows(cur_data_all())], na.rm = TRUE),
    min       = min(proportion_primary_minimum_adj, na.rm = TRUE),
    max       = max(proportion_primary_maximum_adj, na.rm = TRUE),
    mean_min  = mean(proportion_primary_minimum_adj, na.rm = TRUE),
    sd_min    = sd(proportion_primary_minimum_adj, na.rm = TRUE),
    mean_max  = mean(proportion_primary_maximum_adj, na.rm = TRUE),
    sd_max    = sd(proportion_primary_maximum_adj, na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  mutate(
  group_display = if_else(
    grp_composition == "MF", 
    "MF (All)", 
    as.character(grp_composition)  # visible label
  ),
  is_sub = FALSE
)

# 2) MF sub-rows by pairbond_type (only if MF exists)
mf_sub <- data_clean %>%
  filter(grp_composition == "MF") %>%
  mutate(pairbond_type = replace_na(pairbond_type, "Unknown")) %>%
  group_by(grp_composition, pairbond_type) %>%
  summarise(
    n         = sum(valid_rows(cur_data_all())),
    n_species = n_distinct(species_10ktrees[valid_rows(cur_data_all())]),
    n_pats    = sum(n_paternities[valid_rows(cur_data_all())], na.rm = TRUE),
    min       = min(proportion_primary_minimum_adj, na.rm = TRUE),
    max       = max(proportion_primary_maximum_adj, na.rm = TRUE),
    mean_min  = mean(proportion_primary_minimum_adj, na.rm = TRUE),
    sd_min    = sd(proportion_primary_minimum_adj, na.rm = TRUE),
    mean_max  = mean(proportion_primary_maximum_adj, na.rm = TRUE),
    sd_max    = sd(proportion_primary_maximum_adj, na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  mutate(
    # visible label for sub-rows; indent is styled in gt
    group_display = paste0(pairbond_type),
    is_sub = TRUE
  )

# 3) Bind, then order so sub-rows appear directly under MF; keeps original order of grp_comp
order_vec <- main_tbl$grp_composition
# place MF subrows immediately after the MF row
combined <- bind_rows(
  main_tbl,
  mf_sub
) %>%
  mutate(
    grp_order = match(grp_composition, order_vec),
    # make sure MF subrows come right after MF (grp_order + small epsilon)
    sub_bump  = ifelse(is_sub & grp_composition == "MF", 0.5, 0),
    sort_key  = grp_order + sub_bump
  ) %>%
  arrange(sort_key, grp_composition, is_sub, group_display)

# 4) Build gt table (indent MF sub-rows)
# tbl_grptype number is pre-computed in setup chunk as tbl_grptype_num
summary_table <- combined %>%
  select(group_display, n, n_species, n_pats,
         min, max, mean_min, sd_min, mean_max, sd_max, is_sub) %>%
  gt(rowname_col = "group_display") %>%
  tab_stubhead(label = "Group composition") %>%
  tab_header(
    title = paste0("Table ", tbl_grptype_num, ". Proportion of paternities obtained by primary males, by group type")
  ) %>%
  cols_label(
    n               = "N obs",
    n_species       = "N species",
    n_pats          = "Total paternities",
    min             = "Min",
    max             = "Max",
    mean_min        = "Mean (min est.)",
    sd_min          = "SD (min est.)",
    mean_max        = "Mean (max est.)",
    sd_max          = "SD (max est.)"
  ) %>%
  # format numeric columns
  fmt_number(
    columns  = c(min, max, mean_min, sd_min, mean_max, sd_max),
    decimals = 2
  ) %>%
  # left-align numeric columns
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_body(
      columns = c(n, n_species, n_pats, min, max, mean_min, sd_min, mean_max, sd_max)
    )
  ) %>%
  tab_style(
    style     = cell_text(align = "left"),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_stub(rows = !is_sub)
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_stubhead()
  ) %>%
  # indent the stub for sub-rows
  tab_style(
    style = cell_text(indent = px(18)),
    locations = cells_stub(rows = is_sub)
  ) %>%
  # de-emphasize the sub-rows a little
  tab_style(
    style = cell_text(color = "gray30"),
    locations = cells_body(rows = is_sub, columns = everything())
  ) %>%
  tab_options(
    table.border.top.style = "none",
    table.font.size        = px(13),
    table.align            = "left"
  ) %>%
  cols_hide(columns = is_sub)

```


```{r paternity-loss-by-grp-type-model, echo=FALSE, results='hide'} 

#This is basically a fancy ANOVA; only purpose is to determine whether there are meaningful differences in the paternity share of different group types. 

if (file.exists("../output/model-objects/paternityloss_grptype_model.rds")) {
  paternityloss_grptype_model <- readRDS("../output/model-objects/paternityloss_grptype_model.rds")
} else {
paternityloss_grptype_model <- brm(
  bf(proportion_primary_maximum_adj | weights(n_paternities^0.5) ~
       grp_composition_recode +
       (1 | gr(genus_species_subspecies, cov = phylo_matrix)) +
       (1 | species_nonphylo)),
  family = Beta(),
  data = data_clean,
  data2 = list(phylo_matrix = phylo_matrix),
  prior = c(
    prior(normal(0, 1), class = "Intercept"),
    prior(normal(0, 1), class = "b"),
    prior(student_t(3, 0, 2.5), class = "sd"),
    prior(student_t(3, 0, 2.5), class = "phi")
  ),
  chains = 4, iter = 10000, seed = 123,
  control = list(adapt_delta = 0.99)
)
  saveRDS(paternityloss_grptype_model, "../output/model-objects/paternityloss_grptype_model.rds")
}

summary(paternityloss_grptype_model)
#plot(paternityloss_grptype_model)

###Convert to proportion scale for reporting purposes###

# 1. sanity check: levels exist and MM is reference (so no b_*MM column)
stopifnot(all(c("MM","MF","MFF") %in% levels(data_clean$grp_composition_recode)))
dr <- as_draws_df(paternityloss_grptype_model)

# 2. linear predictors (logit) from fixed effects
eta_MM  <- dr$b_Intercept
eta_MF  <- eta_MM + dr$`b_grp_composition_recodeMF`
eta_MFF <- eta_MM + dr$`b_grp_composition_recodeMFF`

# 3. map to response scale (proportions)
mu_MM  <- plogis(eta_MM)
mu_MF  <- plogis(eta_MF)
mu_MFF <- plogis(eta_MFF)

# helper that returns a numeric named vector
summ_vec <- function(x) c(mean = mean(x), lo = unname(quantile(x, 0.025)),
                          hi = unname(quantile(x, 0.975)))

# group means (response scale)
g_MM  <- summ_vec(mu_MM)
g_MF  <- summ_vec(mu_MF)
g_MFF <- summ_vec(mu_MFF)

# pairwise differences (absolute, response scale) — choose direction explicitly
diff_MM_MF_prop  <- mu_MM  - mu_MF
diff_MM_MFF_prop <- mu_MM  - mu_MFF
diff_MF_MFF_prop <- mu_MF  - mu_MFF

d_MM_MF_prop  <- summ_vec(diff_MM_MF_prop)
d_MM_MFF_prop <- summ_vec(diff_MM_MFF_prop)
d_MF_MFF_prop <- summ_vec(diff_MF_MFF_prop)

# 4. hard sanity checks (these will error if something goes off)
# If MFF has higher mean than MM on the logit scale, this should be strictly < 0:
stopifnot(d_MM_MFF_prop["hi"] < 0)

# 5. create fresh names for inline (to avoid collisions with old vars)
MM_mean_prop     <- g_MM["mean"];    MM_lo_prop <- g_MM["lo"];    MM_hi_prop <- g_MM["hi"]
MF_mean_prop     <- g_MF["mean"];    MF_lo_prop <- g_MF["lo"];    MF_hi_prop <- g_MF["hi"]
MFF_mean_prop    <- g_MFF["mean"];   MFF_lo_prop <- g_MFF["lo"];  MFF_hi_prop <- g_MFF["hi"]

# Save Analysis 1 MM mean for use in Question 3 text (before it is overwritten by Analysis 2)
MM_mean_prop_A1 <- g_MM["mean"]
MM_lo_prop_A1   <- g_MM["lo"]
MM_hi_prop_A1   <- g_MM["hi"]
MM_mean_pct_A1_s <- sprintf("%.0f", MM_mean_prop_A1 * 100)
MM_lo_pct_A1_s   <- sprintf("%.0f", MM_lo_prop_A1 * 100)
MM_hi_pct_A1_s   <- sprintf("%.0f", MM_hi_prop_A1 * 100)

MM_MF_diff_prop  <- d_MM_MF_prop["mean"];  MM_MF_lo_prop  <- d_MM_MF_prop["lo"];  MM_MF_hi_prop  <- d_MM_MF_prop["hi"]
MM_MFF_diff_prop <- d_MM_MFF_prop["mean"]; MM_MFF_lo_prop <- d_MM_MFF_prop["lo"]; MM_MFF_hi_prop <- d_MM_MFF_prop["hi"]
MF_MFF_diff_prop <- d_MF_MFF_prop["mean"]; MF_MFF_lo_prop <- d_MF_MFF_prop["lo"]; MF_MFF_hi_prop <- d_MF_MFF_prop["hi"]

# 6. (optional) formatted strings with new names
fmt2 <- function(z) sprintf('%.2f', z)
MM_mean_prop_s     <- fmt2(MM_mean_prop);     MM_lo_prop_s  <- fmt2(MM_lo_prop);   MM_hi_prop_s  <- fmt2(MM_hi_prop)
MF_mean_prop_s     <- fmt2(MF_mean_prop);     MF_lo_prop_s  <- fmt2(MF_lo_prop);   MF_hi_prop_s  <- fmt2(MF_hi_prop)
MFF_mean_prop_s    <- fmt2(MFF_mean_prop);    MFF_lo_prop_s <- fmt2(MFF_lo_prop);  MFF_hi_prop_s <- fmt2(MFF_hi_prop)

MM_MF_diff_prop_s  <- fmt2(MM_MF_diff_prop);  MM_MF_lo_prop_s  <- fmt2(MM_MF_lo_prop);  MM_MF_hi_prop_s  <- fmt2(MM_MF_hi_prop)
MM_MFF_diff_prop_s <- fmt2(MM_MFF_diff_prop); MM_MFF_lo_prop_s <- fmt2(MM_MFF_lo_prop); MM_MFF_hi_prop_s <- fmt2(MM_MFF_hi_prop)
MF_MFF_diff_prop_s <- fmt2(MF_MFF_diff_prop); MF_MFF_lo_prop_s <- fmt2(MF_MFF_lo_prop); MF_MFF_hi_prop_s <- fmt2(MF_MFF_hi_prop)

# Count MMF and MMFF observations for inline reporting
n_mmf  <- sum(data_analysis$grp_composition == "MMF", na.rm = TRUE)
n_mmff <- sum(data_analysis$grp_composition == "MMFF", na.rm = TRUE)

```

```{r data-consistency-check-question-2, echo=FALSE, results='hide', eval=FALSE}

#Verify that the observations being used in the table and model are all the same

#1--Identify the rows the table uses
table_row_ids <- data_clean %>%
  filter(
    !is.na(proportion_primary_minimum_adj) &
    !is.na(proportion_primary_maximum_adj)
  ) %>%
  pull(row_id)

#2--Identify the rows the model uses
model_row_ids <- data_clean %>%
  filter(
    !is.na(proportion_primary_maximum_adj),               # response
    !is.na(n_paternities),                                 # weights
    !is.na(grp_composition_recode),                        # fixed effect
    !is.na(genus_species_subspecies),                      # phylo RE
    !is.na(species_nonphylo)                               # non‐phylo RE
  ) %>%
  pull(row_id)

#3--Which rows are in the table but not in the model?
rows_only_in_table <- setdiff(table_row_ids, model_row_ids)

#4--Which rows are in the model but not in the table?
rows_only_in_model <- setdiff(model_row_ids, table_row_ids)

#If both rows_only objects are empty, observations match

```

__Analysis 1: Description of relationship between paternity share and group composition__

We first evaluated the relationship between group composition and paternity share in the raw data (Table `r tbl_grptype_num`). On average, primary males obtained `r mean_val_primary_s`% (SD = `r sd_val_primary_s`%) of paternities across all group types; we obtained a similar value when all uncertain paternities were assigned to other males (mean = `r mean_val_primary_min_s`%, SD = `r sd_val_primary_min_s`%). 

For this analysis, we grouped the multi-male/single-female observations (n=`r n_mmf`) and multi-male/multi-female observations (n=`r n_mmff`) into a single multi-male category. These single-female groups could not be analyzed as a separate category due to the small sample size, and both observations had primary male paternity share proportions that were well within the range for multi-male/multi-female groups (Table `r tbl_grptype_num`, Figure `r fig_patloss`). 

Primary males in single-male/multi-female groups had the highest average paternity share (posterior mean=`r MFF_mean_prop_s`, 95% CI: `r MFF_lo_prop_s` - `r MFF_hi_prop_s`) and those in multi-male groups the lowest average share (posterior mean=`r MM_mean_prop_s`, 95% CI: `r MM_lo_prop_s` - `r MM_hi_prop_s`), though the variation within each category was considerable (Table `r tbl_grptype_num`, Figure `r fig_patloss`). The difference between single-male/multi-female and multi-male groups was credible (95% CI: `r MM_MFF_lo_prop_s` - `r MM_MFF_hi_prop_s`), while the contrast between pair-living and multi-male groups was weaker, with a 95% CI that barely crossed zero (`r MM_MF_lo_prop_s` - `r MM_MF_hi_prop_s`). The unexpectedly low paternity share in pair-living species was driven by dispersed pairs rather than cohesive pairs (Figure `r fig_patloss`).

```{r summary_table, echo=FALSE}

summary_table

```

```{r fig3-paternity-loss-model-plot, echo=FALSE, results='hide', fig.cap=fig_patloss_cap}

# Get conditional effects from the model
pl_ce <- conditional_effects(paternityloss_grptype_model, "grp_composition_recode")
pl_ce_data <- pl_ce$grp_composition_recode

# Plot conditional effects with raw data, colored by MF pairbond type

# 1. Create a new column to classify for color
data_clean <- data_clean %>%
  mutate(plot_color = case_when(
    grp_composition_recode == "MF" & pairbond_type == "cohesive" ~ "Cohesive MF",
    grp_composition_recode == "MF" & pairbond_type == "dispersed" ~ "Dispersed MF",
    grp_composition == "MMF" ~ "MMF",
    TRUE ~ "Other"
  ))

# 2. Compute raw means for the two MF subtypes (for diamond + label)
mf_means <- data_clean %>%
  filter(grp_composition_recode == "MF", !is.na(pairbond_type)) %>%
  group_by(grp_composition_recode, plot_color) %>%
  summarise(
    mean_val = mean(proportion_primary_maximum_adj, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(label = if_else(plot_color == "Cohesive MF", "cohesive", "dispersed"))

# 3. Define custom colors
mf_colors <- c(
  "Cohesive MF" = "#1b9e77",   # green
  "Dispersed MF" = "#d95f02",  # orange
  "MMF" = "#2196F3",           # blue
  "Other" = "gray60"
)

# 4. Compute mean position for MMF label
mmf_means <- data_clean %>%
  filter(plot_color == "MMF") %>%
  summarise(
    grp_composition_recode = first(grp_composition_recode),
    mean_val = mean(proportion_primary_maximum_adj, na.rm = TRUE),
    .groups = "drop"
  )

# 4. Plot
ggplot(pl_ce_data, aes(x = effect1__, y = estimate__)) +
  # Raw data points
  geom_jitter(data = data_clean,
              aes(x = grp_composition_recode, y = proportion_primary_maximum_adj, color = plot_color),
              width = 0.1, height = 0, alpha = 0.5, size = 1.5) +

  # Posterior means and CIs
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = lower__, ymax = upper__), width = 0.2, linewidth = 0.8) +

  # Colored diamonds for MF subtypes
  stat_summary(
    data = data_clean %>% filter(plot_color %in% c("Cohesive MF", "Dispersed MF")),
    aes(x = grp_composition_recode, y = proportion_primary_maximum_adj, color = plot_color),
    fun = mean,
    geom = "point",
    shape = 18,  # diamond
    size = 4
  ) +

  # Labels for each MF subtype mean (offset for visibility)
geom_text(
  data = mf_means,
  aes(x = grp_composition_recode, y = mean_val, label = label, color = plot_color),
  size = 4,
  fontface = "italic",
  hjust = 0,
  nudge_x = 0.15,
  nudge_y = -0.02
) +

  # Label for MMF observations
geom_text(
  data = mmf_means,
  aes(x = grp_composition_recode, y = mean_val, label = "multi-male/\nsingle-female"),
  color = "#2196F3",
  size = 4,
  fontface = "italic",
  hjust = 0,
  nudge_x = 0.12,
  nudge_y = -0.02
) +

  scale_color_manual(values = mf_colors) +
  guides(color = "none") +  # remove legend

  labs(
    x = "Group composition",
    y = "Estimated primary male paternity share",
    title = "Predicted paternity share by group type"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 10)),
    plot.title = element_text(hjust = 0.5)
  )




```

```{r competition-monopolizability-model, echo=FALSE, results='hide', fig.show='hide'}

if (file.exists("../output/model-objects/competition_monopolizability_model.rds")) {
  competition_monopolizability_model <- readRDS("../output/model-objects/competition_monopolizability_model.rds")
} else {
  competition_monopolizability_model <- brm(
    formula = bf(
      proportion_primary_maximum_adj | weights(n_paternities^0.5) ~ 
        n + m + m2 + n:m +
        (1 | gr(genus_species_subspecies, cov = phylo_matrix)) + 
        (1 | species_nonphylo)
    ),
    data = data_clean,
    family = Beta(link = "logit"),
    data2 = list(phylo_matrix = phylo_matrix),
    prior = c(
      prior(normal(0, 1), class = "Intercept"),
      prior(normal(0, 1), class = "b"),
      prior(student_t(3, 0, 2.5), class = "sd"),
      prior(student_t(3, 0, 2.5), class = "phi")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4, cores = 4, iter = 10000, seed = 123
  )
  saveRDS(competition_monopolizability_model, "../output/model-objects/competition_monopolizability_model.rds")
}

summary(competition_monopolizability_model)
```

```{r extract-means-CIs-competition-monopolizability-model, echo=FALSE, results='hide', fig.show='hide'}

#####Extract means and CIs for competition var; how big is the diff between multi and single male groups?######

# Extract posterior draws
dr <- as_draws_df(competition_monopolizability_model)

# Compute linear predictors for SM and MM group type, holding m/m^2 constant
# Choose mean m (seasonality)
m_val <- mean(data_clean$m, na.rm = TRUE)
m2_val <- m_val^2

# Group 1: MM (n = 0) — reference group
eta_MM <- dr$b_Intercept + 
          dr$b_m   * m_val + 
          dr$b_m2  * m2_val

# Group 2: SM (n = 1)
eta_SM <- dr$b_Intercept + 
          dr$b_n   + 
          dr$b_m   * m_val + 
          dr$b_m2  * m2_val + 
          dr$`b_n:m` * m_val

# Transform to response scale (inverse logit)
mu_MM <- plogis(eta_MM)
mu_SM <- plogis(eta_SM)

# Summarize posterior for each group (reusing summ_vec from earlier)
g_MM <- summ_vec(mu_MM)
g_SM <- summ_vec(mu_SM)

# Format for use in text (reusing fmt2 from earlier)

MM_mean_prop_s <- fmt2(g_MM["mean"])
MM_lo_prop_s   <- fmt2(g_MM["lo"])
MM_hi_prop_s   <- fmt2(g_MM["hi"])

SM_mean_prop_s <- fmt2(g_SM["mean"])
SM_lo_prop_s   <- fmt2(g_SM["lo"])
SM_hi_prop_s   <- fmt2(g_SM["hi"])

# Calculate difference in group types across posterior draws
diff_SM_MM <- mu_SM - mu_MM

#Summary of the group differnce (mean/CI)
d_SM_MM <- summ_vec(diff_SM_MM)
d_SM_MM

#Format for use in text
SM_MM_diff_prop_s  <- fmt2(d_SM_MM["mean"])
SM_MM_lo_prop_s    <- fmt2(d_SM_MM["lo"])
SM_MM_hi_prop_s    <- fmt2(d_SM_MM["hi"])

# Sample size for this model (smaller than Analysis 1 b/c not all species have seasonality data)
n_obs_cm <- nobs(competition_monopolizability_model)
n_species_cm <- length(unique(
  sub("^([^_]+_[^_]+).*", "\\1", competition_monopolizability_model$data$genus_species_subspecies)
))
n_pats_cm <- sum(competition_monopolizability_model$data$n_paternities, na.rm = TRUE)

n_obs_cm_fmt <- format(n_obs_cm, big.mark = ",")
n_species_cm_fmt <- format(n_species_cm, big.mark = ",")
n_pats_cm_fmt <- format(n_pats_cm, big.mark = ",")

```

```{r plot-competition-monopolizability-model, echo=FALSE, results='hide', fig.show='hide'}

#################Plot competition monopolizability model#############

# Choose m levels 
m_vals <- quantile(data_clean$m, probs = c(0.1, 0.5, 0.9), na.rm = TRUE)

newdat_B <- expand.grid(
  n = c(0, 1),
  m = as.numeric(m_vals)
) %>%
  mutate(
    m2    = m^2,
    n_lab = factor(n, levels = c(0,1), labels = c("MM", "SM")),
    m_lab = paste0("m = ", round(m, 2))
  )

# Get summaries and rename to mean/lo/hi
pred_B <- fitted(
  competition_monopolizability_model,
  newdata    = newdat_B,
  re_formula = NA,
  summary    = TRUE,
  scale      = "response"
) %>%
  as_tibble() %>%
  transmute(mean = Estimate, lo = `Q2.5`, hi = `Q97.5`)

summ_B <- bind_cols(newdat_B, pred_B)

# Sanity check (should include mean, lo, hi)
# print(names(summ_B))

# Order legend by increasing m
levels_m <- paste0("m = ", round(sort(unique(summ_B$m)), 2))
summ_B <- summ_B %>%
  mutate(m_lab = factor(m_lab, levels = levels_m))

# Color-blind friendly palette for figure
okabe_ito <- c("#56B4E9", "#009E73", "#E69F00", "#0072B2",
               "#D55E00", "#CC79A7", "#F0E442", "#000000")

pd <- position_dodge(width = 0.35)
brks <- levels(summ_B$m_lab)

cm_plot <- ggplot(summ_B, aes(x = n_lab, y = mean, ymin = lo, ymax = hi, color = m_lab)) +
  geom_point(position = pd, size = 2.8, shape = 16) +   
  geom_errorbar(position = pd, width = 0.12, linewidth = 0.6) +
  scale_color_manual(values = okabe_ito, name = "Monopolizability (m)") +
  labs(x = "Group composition", y = "Estimated primary male paternity share") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right") +
  scale_color_manual(
    breaks = brks,
    labels = c("most seasonal", "moderately seasonal", "least seasonal"),
    values = okabe_ito[seq_along(brks)],  
    name   = "Seasonality")

```

```{r derivatives-competition-monopolizability-model, echo=FALSE, results='hide', fig.show='hide'}

######Extract derivatives; as monopolizability increases, how fast does paternity share change?

# Partial derivative of logit(mu) w/r/t monopolizability at the mean
m_deriv <- mean(data_clean$m, na.rm = TRUE)

# Logit-scale derivative: d logit(mu) / dm = b_m + 2*b_m2*m + b_n:m * n
dlogit_MM <- dr$b_m + 2 * dr$b_m2 * m_deriv                    # n = 0
dlogit_SM <- dr$b_m + 2 * dr$b_m2 * m_deriv + dr$`b_n:m`       # n = 1

# Response-scale derivative via chain rule: d mu / dm = dlogit/dm * mu * (1 - mu)
# Reuse eta and mu from earlier chunk (eta_MM, eta_SM, mu_MM, mu_SM)
dresp_MM <- dlogit_MM * mu_MM * (1 - mu_MM)
dresp_SM <- dlogit_SM * mu_SM * (1 - mu_SM)

# Summarize
summ_vec(dlogit_MM)
summ_vec(dlogit_SM)
summ_vec(dresp_MM)
summ_vec(dresp_SM)

# Format for text
dresp_MM_mean <- fmt2(summ_vec(dresp_MM)["mean"])
dresp_MM_lo   <- fmt2(summ_vec(dresp_MM)["lo"])
dresp_MM_hi   <- fmt2(summ_vec(dresp_MM)["hi"])

dresp_SM_mean <- fmt2(summ_vec(dresp_SM)["mean"])
dresp_SM_lo   <- fmt2(summ_vec(dresp_SM)["lo"])
dresp_SM_hi   <- fmt2(summ_vec(dresp_SM)["hi"])

```

__Analysis 2: Evaluation of the presence of in-group competitors and seasonality__

We next examined how the presence of both in-group competitors and seasonality influenced the paternity share of primary males. Seasonality data were not available for all species, so this analysis was based on a slightly smaller sample than Analysis 1 (`r n_obs_cm_fmt` observations, `r n_species_cm_fmt` species, and `r n_pats_cm_fmt` total paternities). In these models, both types of single-male groups (i.e., pair-living and single-male/multi-female) were combined into one category called single-male, because Analysis 1 estimates of primary males’ paternity shares in these groups were not reliably different and because they share the same key characteristic of having no in-group competitors. 

The extent of reproductive seasonality had little effect on primary male paternity share, regardless of whether males lived in single or multi-male groups (Figure `r fig_seasonality`). Consistent with this overall conclusion, the estimated change in paternity share per unit increase in monopolizability was near zero in both multi-male groups (posterior mean = `r dresp_MM_mean`, 95% CI: `r dresp_MM_lo` - `r dresp_MM_hi`) and in single-male groups (posterior mean = `r dresp_SM_mean`, 95% CI: `r dresp_SM_lo` - `r dresp_SM_hi`). The results were qualitatively very similar when pair-living groups, where mating opportunities should be highly monopolizable regardless of seasonality, were excluded. 

Adding seasonality to the model slightly lowered the expected paternity share for primary males (Figure `r fig_seasonality`). Here, for multi-male groups the posterior mean was `r MM_mean_prop_s`, while for single-male groups it was `r SM_mean_prop_s`. The 95% CI of the difference between the two did not include zero (95% CI: `r SM_MM_lo_prop_s` - `r SM_MM_hi_prop_s`), indicating a credible difference. 

```{r competition-monopolizability-plot, echo=FALSE, fig.cap=fig_seasonality_cap}

cm_plot

```

**Question 3: When primary males in multimale groups lose paternity, who are they losing it to?**

```{r egp-proportions-nums, echo=FALSE, results='hide'}

# Raw summary numbers — need this in order to fill in inline text numbers plus table

# A single “raw” summary with full precision
summary_data_egps_raw <- data_clean %>%
  filter(grp_composition_recode == "MM") %>%
  filter(
    !is.na(proportion_egp_minimum_adj),
    !is.na(proportion_egp_maximum_adj)
  ) %>%
  summarise(
    n_obs     = n(),                                            # obs
    n_pats    = sum(n_paternities, na.rm = TRUE),               # total paternities
    n_species = n_distinct(species_10ktrees),           # species
    min       = min(proportion_egp_minimum_adj, na.rm = TRUE), # raw min prop
    max       = max(proportion_egp_maximum_adj, na.rm = TRUE), # raw max prop
    mean_min  = mean(proportion_egp_minimum_adj, na.rm = TRUE), # raw mean(min)
    sd_min    = sd(proportion_egp_minimum_adj, na.rm = TRUE),   # raw sd(min)
    mean_max  = mean(proportion_egp_maximum_adj, na.rm = TRUE), # raw mean(max)
    sd_max    = sd(proportion_egp_maximum_adj, na.rm = TRUE)    # raw sd(max)
  )

# Reformatting numbers for use in text
mean_min_str <- sprintf("%.0f", summary_data_egps_raw$mean_min * 100)
sd_min_str   <- sprintf("%.0f", summary_data_egps_raw$sd_min   * 100)
mean_max_str <- sprintf("%.0f", summary_data_egps_raw$mean_max * 100)
sd_max_str   <- sprintf("%.0f", summary_data_egps_raw$sd_max   * 100)
obs_str      <- format(summary_data_egps_raw$n_obs,     big.mark = ",")
species_str  <- format(summary_data_egps_raw$n_species, big.mark = ",")
pats_str     <- format(summary_data_egps_raw$n_pats,    big.mark = ",")

# EGP share of "lost" paternities (using Analysis 1 MM posterior mean)
egp_loss_share_min <- summary_data_egps_raw$mean_min / (1 - MM_mean_prop_A1)
egp_loss_share_max <- summary_data_egps_raw$mean_max / (1 - MM_mean_prop_A1)
egp_loss_share_min_str <- sprintf("%.0f", egp_loss_share_min * 100)
egp_loss_share_max_str <- sprintf("%.0f", egp_loss_share_max * 100)

```





```{r troubleshooting-n-differences, echo=FALSE, results='hide', eval=FALSE}

#Figure out why I get more EGP values than primary male values for MMFF groups.

# Count MMF and MMFF rows with different missingness profiles
data_clean %>%
  filter(grp_composition_recode == "MM") %>%
  mutate(
    has_primary = !is.na(proportion_primary_minimum_adj) & !is.na(proportion_primary_maximum_adj),
    has_egp = !is.na(proportion_egp_minimum_adj) & !is.na(proportion_egp_maximum_adj)
  ) %>%
  count(grp_composition, has_primary, has_egp)

# Identify rows where EGP data exist but primary paternity data do not
egp_only_cases <- data_clean %>%
  filter(
    !is.na(proportion_egp_minimum_adj) & !is.na(proportion_egp_maximum_adj) & 
    (is.na(proportion_primary_minimum_adj) | is.na(proportion_primary_maximum_adj))
  )

# Select relevant identifying information
egp_only_cases_summary <- egp_only_cases %>%
  select(genus_species_subspecies, reference, grp_composition) %>%
  distinct()

egp_only_cases_summary

```

```{r paternityloss-source-model, echo=FALSE, results='hide', fig.show='hide'}

loss_source_data_mm <- data_subset_mm %>%
  mutate(
    # Calculate in-group losses and reformat negatives
    n_igp        = as.integer(n_paternities - n_alphaprimary - n_egp),
    n_igp        = pmax(n_igp, 0L),
    # Total losses
    total_losses = n_egp + n_igp,
    # Prepare weights
    model_weights = n_paternities^0.5
  ) %>%
  filter(
    # Keep zero-loss cases (total_losses >= 0)
    !is.na(n_egp),
    !is.na(n_igp),
    !is.na(n_paternities),
    !is.na(genus_species_subspecies),
    !is.na(species_nonphylo)
  )


if (file.exists("../output/model-objects/loss_source_model_mm.rds")) {
  loss_source_model_mm <- readRDS("../output/model-objects/loss_source_model_mm.rds")
} else {
  loss_source_model_mm <- brm(
  bf(n_egp | trials(total_losses) ~
       (1 | gr(genus_species_subspecies, cov = phylo_matrix)) +
       (1 | species_nonphylo)),
  family = binomial(),
  data = loss_source_data_mm,
  data2 = list(phylo_matrix = phylo_matrix),
  prior = c(
    prior(normal(0, 1), class = "Intercept"),
    prior(student_t(3, 0, 2.5), class = "sd")
  ),
  sample_prior = "no",
  chains = 4, iter = 10000, seed = 123,
  control = list(adapt_delta = 0.99)
)
  saveRDS(loss_source_model_mm, "../output/model-objects/loss_source_model_mm.rds")
}

summary(loss_source_model_mm)

# Use posterior_linpred (not posterior_epred) to get the probability (pi)
# on the response scale. posterior_epred returns expected counts for binomial
# models (n * pi), which is not what we want here.
egp_loss_summary <- posterior_linpred(loss_source_model_mm, transform = TRUE, re_formula = NA) %>%
  as.data.frame() %>%
  summarise(
    mean = round(mean(V1) * 100, 0),
    ci_lower = round(quantile(V1, 0.025) * 100, 0),
    ci_upper = round(quantile(V1, 0.975) * 100, 0)
  )

egp_loss_summary

igp_loss_summary <- posterior_linpred(loss_source_model_mm, transform = TRUE, re_formula = NA) %>%
  as.data.frame() %>%
  summarise(
    mean = round(mean(1 - V1) * 100, 0),
    ci_lower = round(quantile(1 - V1, 0.025) * 100, 0),
    ci_upper = round(quantile(1 - V1, 0.975) * 100, 0)
  )

igp_loss_summary

n_obs_loss_source <- nrow(loss_source_model_mm$data)
n_species_loss_source <- length(unique(
  sub("^([^_]+_[^_]+).*", "\\1", loss_source_model_mm$data$genus_species_subspecies)
))
n_pats_loss_source <- sum(loss_source_data_mm$n_paternities, na.rm = TRUE)

n_obs_loss_source_fmt <- format(n_obs_loss_source, big.mark = ",")
n_species_loss_source_fmt <- format(n_species_loss_source, big.mark = ",")
n_pats_loss_source_fmt <- format(n_pats_loss_source, big.mark = ",")

# Derived quantities for inline text: overall EGP rate based on Analysis 1 CI and model EGP share
loss_hi_A1 <- 1 - MM_lo_prop_A1   # max share lost (from low end of CI)
loss_lo_A1 <- 1 - MM_hi_prop_A1   # min share lost (from high end of CI)
loss_lo_pct_s <- sprintf("%.0f", loss_lo_A1 * 100)
loss_hi_pct_s <- sprintf("%.0f", loss_hi_A1 * 100)

egp_frac <- egp_loss_summary$mean / 100   # model-estimated EGP proportion of losses
overall_egp_lo <- egp_frac * loss_lo_A1
overall_egp_hi <- egp_frac * loss_hi_A1
overall_egp_lo_s <- sprintf("%.0f", overall_egp_lo * 100)
overall_egp_hi_s <- sprintf("%.0f", overall_egp_hi * 100)

pp_check(loss_source_model_mm, type = "hist")
pp_check(loss_source_model_mm, type = "dens_overlay")

```

```{r loss-source-dropped-obs-diagnostic, echo=FALSE, results='hide', eval=FALSE}

#Determine why a bunch of obs drop from the model

#1 Rows used in the raw summary
raw_ids <- data_clean %>%
  filter(
    grp_composition_recode == "MM",
    !is.na(proportion_egp_minimum_adj),
    !is.na(proportion_egp_maximum_adj)
  ) %>%
  pull(row_id)

#2 Rows used in the data for model 
model_ids <- loss_source_data_mm$row_id

#3 Which rows are in the raw summary but not in the model?
dropped_ids <- setdiff(raw_ids, model_ids)

#4 Look at the dropped rows
dropped_obs <- data_clean %>%
  filter(row_id %in% dropped_ids) %>%
  select(
    row_id,
    grp_composition_recode,
    proportion_egp_minimum_adj,
    proportion_egp_maximum_adj,
    n_paternities,
    n_egp,
    n_alphaprimary,
    genus_species_subspecies,
    species_nonphylo
  )

#5 Print them all so I can see exactly why each was dropped // These are rows that do not have n_primaryalpha, which is necessary to calculate losses.

#print(dropped_obs, n = Inf)

```

In the raw data, a minimum of `r mean_min_str`% (SD = `r sd_min_str`%) of paternities in multi-male groups were assigned to extra-group males; this rose to `r mean_max_str`% (SD = `r sd_max_str`%) if all uncertain paternities were assigned to extra‐group males (n=`r obs_str` observations, `r species_str` species, and `r pats_str` total paternities). If primary males in multi-male groups sire approximately `r MM_mean_pct_A1_s`% of infants (i.e., the posterior mean prediction from Question 2, Analysis 1 above), then about `r sprintf("%.0f", (1 - MM_mean_prop_A1) * 100)`% of infants are sired by other males. Given a raw, overall EGP rate of `r mean_min_str`–`r mean_max_str`%, this would mean that `r egp_loss_share_min_str`–`r egp_loss_share_max_str`% of the paternities that primary males lose are attributable to extra-group males, with the remainder lost to group mates.^[`r mean_min_str`/`r sprintf("%.0f", (1 - MM_mean_prop_A1) * 100)` = `r egp_loss_share_min_str`%; `r mean_max_str`/`r sprintf("%.0f", (1 - MM_mean_prop_A1) * 100)` = `r egp_loss_share_max_str`%.]

While our statistical models generally tell a similar story, they suggest that losses to extra-group males may be slightly higher than values obtained from the raw data. For the subset of data where all relevant information was available (`r n_obs_loss_source_fmt` observations, `r n_species_loss_source_fmt` species, and `r n_pats_loss_source_fmt` paternities), an estimated `r egp_loss_summary$mean`% of primary males' lost paternities were EGPs (95% CI: `r egp_loss_summary$ci_lower` - `r egp_loss_summary$ci_upper`%). The other `r igp_loss_summary$mean`% of lost paternities went to in-group males (95% CI: `r igp_loss_summary$ci_lower` - `r igp_loss_summary$ci_upper`%). 

Since the 95% CI for primary male paternity share in multimale groups is `r MM_lo_pct_A1_s`-`r MM_hi_pct_A1_s`% (see Figure `r fig_patloss`), this means that somewhere between `r loss_lo_pct_s`% and `r loss_hi_pct_s`% of infants are sired by other males. If `r egp_loss_summary$mean`% of the low end and high end of these paternity share estimates are EGPs, this would translate to an overall EGP rate of ~`r overall_egp_lo_s`-`r overall_egp_hi_s`%.^[`r sprintf("%.2f", egp_frac)` × `r sprintf("%.2f", loss_lo_A1)` = `r sprintf("%.2f", overall_egp_lo)`; `r sprintf("%.2f", egp_frac)` × `r sprintf("%.2f", loss_hi_A1)` = `r sprintf("%.2f", overall_egp_hi)`.]    

## Discussion

One of the primary goals of this project was to compile all available genetic paternity data for wild nonhuman primates and produce a single publicly accessible repository for them. The database that we have created contains information about `r n_species_total` species from `r n_genera` genera representing `r n_families` of 14 primate families in 10kTrees. The three families that are not included (*Galagidae*, *Lorisidae*, and *Daubentonidae*) are all largely solitary Strepsirrhines. Although these numbers are gratifyingly large, we are missing information from a substantial number of species, while some families, particularly the *Cercopithecidae* and *Hominidae*, are over-represented in the dataset. 

Phylogeny has a detectable but moderate effect on the distribution of paternity in primate groups. Species-level random effects, which do not differentiate between phylogenetic and non-phylogenetic species differences, explained roughly 35–40% of the total variation in primary males’ paternity share. Our models indicate that between `r round(summary_lambda_iom_min[2] * 100)`% and `r round(summary_lambda_iom[2] * 100)`% of that inter-species variation is attributable to phylogenetic relationships. In absolute terms, this means phylogeny accounts for a relatively modest share of total variation in paternity outcomes. Instead, most of the variation came from observation-level differences within and between species that are not captured by the phylogenetic or species-level random effects. @KamilarCooper2013 found a strong phylogenetic signal associated with morphological traits related to the intensity of intrasexual selection on males, including sexual dimorphism in body mass ($\lambda$ = 0.94), canine size ($\lambda$ = 0.82), and testes size ($\lambda$ = 0.89). They also reported a Pagel’s $\lambda$ of 0.60 for the number of males in groups. Because our $\lambda$ is computed differently than Pagel’s $\lambda$ (i.e., on the random effects derived from our statistical model, rather than directly from the phylogenetic tree), we can't directly compare our estimates to those values. However, the general pattern is consistent: morphological traits associated with male-male competition show very strong phylogenetic signal, while behavioral and demographic traits like group composition and paternity share show comparatively weaker signals.

### Primary males’ paternity share is strongly influenced by the kinds of groups in which they live

Interestingly, pair-living non-human primate species encompass both lowest and highest rates of paternity loss for primary males. For species that live in cohesive pairs, primary males sire about 90% of their partners’ offspring. For species that live in dispersed pairs, primary males sire only about 55%. Similar findings have been reported for a broader range of pair-living mammalian species [@Clutton-BrockIsvaran2006; @Hucketal2014]. @CohasAllaine2009 found no difference between pair-living species that form dispersed and cohesive pair bonds, but they reported that solitary species have much higher rates of extra group paternities than pair-living species. However, they categorized as solitary at least one species, the fork-marked lemur (*Phaner furcifer*), that other researchers have described as having dispersed pair bonds [@SchulkeKappeler2003]. Thus, the preponderance of evidence suggests that for pair-living mammals, paternity patterns are strongly influenced by patterns of association. This highlights the problems with assuming that pair-living (or social monogamy) is a homogenous phenomena, a point which has been made before by other authors [e.g., @Fernandez-Duqueetal2020].

Males in single-male/multi-female groups sire the great majority of offspring within their groups, but do not monopolize paternity as successfully as males in cohesive pairs do. This is likely to reflect the fact that competition over access to groups of females is more intense than competition over access to single females and it is harder for males to monopolize multiple females than to monopolize single females. Blue monkeys, for example, typically form single-male/multi-female groups, but there are regular incursions by outside males during the mating season. Larger numbers of sexually receptive females draw larger numbers of outside males [@GaoCords2020]. 

Primary males living in multi-male groups lose more paternities than primary males living in cohesive pairs or in single-male/multi-female groups. @Clutton-BrockIsvaran2006 reported similar patterns. Paternity losses by primary males in multi-male groups may be the product of several different mechanisms. First, primary males may not be able to monopolize paternity when more than one sexually receptive female is present [@Altmann1962]. Overlap in female receptivity is particularly likely to occur in multi-male/multi-female groups because there are larger numbers of females in these groups [@Ostneretal2008]. Second, in species with long male tenures, females may avoid mating with closely related males. The tenures of primary males in white faced capuchin groups are sometimes long enough that females live in the same groups as their fathers, and males don’t sire the offspring of closely related females [@Godoyetal2016; @Wikbergetal2017]. Third, males in multi-male groups may form coalitions that influence their access to sexually receptive females. Middle-ranking male olive and yellow baboons team up to disrupt consortships of high-ranking males [@Packer1977; @Noe1992], although we don’t know whether these consort takeovers influence paternity. Male chimpanzees rely on their social ties with other males to obtain and maintain high rank and to defend the boundaries of their territories. Alpha males selectively tolerate matings by their allies [@Brayetal2016; @Duffyetal2007], and this translates into paternity success for males with close ties to the alpha male [@Feldblumetal2021].

### Primary males in multi-male groups mostly lose paternities to other resident males 

Our data also show that primary males in multi-male groups lose the majority of paternities to other males within their groups, not to extra-group males. This may be because resident males actively deter incursions by outside males or because subordinate residents are better able to capitalize on reproductive opportunities than outside males. For example, male chacma baboons “monitor consortships so assiduously that they rapidly recognize both when an unexpected mating opportunity arises and when a consortship has ended“ [@Crockfordetal2007, p. 889]. This may give them an important advantage over extra-group males. 

### Seasonality has little effect on primary males' paternity share

Contrary to our predictions, the extent of reproductive seasonality had little effect on the paternity share of primary males.  This prediction was based on the assumption that when reproductive seasonality was strong, there would be more extensive overlap in females’ sexual receptivity, and this would limit primary males’ ability to monopolize conceptions.  In retrospect, we should have been more skeptical of this assumption. @GogartenKoenig2013 showed that male reproductive skew declines as the extent of reproductive synchrony among females increases in non-human primate groups, but the extent of reproductive synchrony among females does not vary consistently with the extent of reproductive seasonality. 


### Limitations of the study

Our study has a number of limitations. First, almost all of the samples are incomplete. In some cases, not all infants born in the population during the study period were sampled, not all potential sires were sampled, or it was not possible to exclude more than one male as the sire of a particular infant. We have estimated the minimum and maximum values of the primary males’ paternity share and the proportion of extra-group paternities, but the real values lie somewhere in between. Second, most species are represented by a single study population, and sometimes a single group. This may bias estimates of the range of variation within and between species (see Alberts et al. XXXX). Third, many species are missing from our database and certain taxa are over-represented.  

In addition, the exercise of writing out our theoretical model made it clear that there is a significant disconnect between the data that would ideally be available to answer our questions, versus the data that are actually available. In particular, it would be best to use conception-level data, to have quantitative information about the degree of extra-group competition that males face, and to have counts of the number of other sexually receptive females present during the probable window of conception. We acknowledge that these are challenging data to collect, especially in species with concealed ovulation. 

### Conclusions

In many respects, our analysis of the largest genetic paternity set assembled on primates to date confirms results from other studies. Primary males consistently obtain the largest share of paternities, but their odds are lower when they live in groups with other males than when they do not. It is also clear that despite this general pattern, there is considerable variation in primary males' ability to control reproduction in their groups, regardless of social group structure. 

However, our results also contained some surprises. Breeding seasonality, which in theory should be a proxy for males' ability to monopolize mating opportunities, did not predict primary males' paternity share regardless of the type of group they lived in (though note that prior studies have found mixed evidence both for effects of seasonality itself, and for the idea that seasonality necessarily means high reproductive synchrony: @KutsukakeNunn2006; @Ostneretal2008; @Paul1997; @GogartenKoenig2013). In addition, pair-living species had both the highest and lowest rates of primary male paternity share. We hope that the availability of our living database will help other researchers 

# Acknowledgements

We would like to thank Adrian Jaggi, Eduardo Fernandez-Duque, and Jason Kamilar for their advice and thoughts on these analyses, as well as Stephanie Fox and Pascale Sicotte, Carola Borries, Sylvain Gatti, Liza Moscovice and Robert Seyfarth, and [XXXX whoever else Joan wants to thank for the olive baboon data] for contributing unpublished paternity. Maud Mouginot, Amy Scott, Joseph Feldblum, Songtao Guo, Caroline Nievergelt, Carola Borries, and [who am I forgetting? XXXX]  all helped us untangle and extract data from published papers, for which we are extremely grateful. 

# References